### 2025-05-20 doc/paper/main.tex
temp

\documentclass[11pt,a4paper]{article}
\usepackage[left=2cm,right=2cm]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}

\title{MPX: Mixed Precision Training for JAX}
\author{}
\date{}

\begin{document}

\maketitle

\section{Introduction}
This paper presents \textbf{mpx}, a tool for training JAX models using mixed precision. The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{equinox}. By leveraging Equinox's flexibility, mpx provides a solution that works with any PyTree structure.

\section{Basics of Mixed Precision Training}
This section summarizes the original Mixed Precision method from NVIDIA's Automatic Mixed Precision \cite{nvidia_amp} and the paper by Micikevicius et al. \cite{mixed_precision_paper}.

Mixed Precision training involves performing most computations in the forward and backward passes of a neural network using 16-bit floating-point numbers. This approach offers two main advantages:
\begin{itemize}
    \item Reduced GPU memory usage (approximately 50\% compared to full precision)
    \item Potential speedup through decreased memory access times and utilization of specialized half-precision tensor cores of TPUs or GPUs.
\end{itemize}

\subsection{Loss Scaling}
A critical component of successful Mixed Precision training is loss scaling. Due to float16's decreased resolution, small gradients may be cast to zero, negatively impacting training performance. The solution involves:
\begin{enumerate}
    \item Scaling the loss by a factor $> 1$
    \item Calculating gradients with the scaled loss
    \item Converting gradients to float32
    \item Dividing by the scaling factor to obtain original gradients
\end{enumerate}

The scaling factor can be automatically adjusted using a simple heuristic:
\begin{itemize}
    \item If scaled gradients exceed float16 range (inf), reduce scaling and skip model update
    \item If scaled gradients remain within float16 range for an extended period, increase scaling
\end{itemize}

\section{Implementation Details}
mpx is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
For this, it provides several functions that allow to cast

The mpx library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:

\begin{enumerate}
    \item \textbf{Transformations to Cast PyTrees}: mpx features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}. All these functions cast all leaves of the input that are JAX arrays and of type float to the corresponding float datatype. All other leaves, including arrays that are of non-float types, like int32, remain unchanged. 
    \item \textbf{Transformations to Cast Functions}: mpx contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions. This transformation returns a function that casts all its inputs to the desired input datatype (using \texttt{cast\_tree(tree, dtype)}), calls the function and then casts the outputs of the function. Moreover, mpx contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision. This is important as some operations, such as sum, mean or softmax, are sensitive to overflows when calculated in float16.
    \item \textbf{Transformations to Cast Gradients}: mpx contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above). Additional to calculating the gradient, the functions also perform the automatic adaption of the loss scaling value.
    These drop-in replacements allow users to reuse their existing Equinox training pipelines without major changes to their structure (cf. Section todo).
\end{enumerate}

\subsection{Automatic Loss Scaling Implementation}
The \texttt{DynamicLossScaling} class manages dynamic loss scaling with methods:
\begin{itemize}
    \item \texttt{scale(x)}: Apply current loss scaling factor
    \item \texttt{unscale(x)}: Remove loss scaling factor
    \item \texttt{adjust(grads\_finite)}: Update scaling factor based on gradient stability
\end{itemize}

\subsection{Optimizer}
mpx works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
The  \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} function handles model updates based on gradient finiteness.
This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{mpx.optimizer\_update}.

\section{Example}
I t

\section{Acknowledgements}
We express our gratitude to Patrick Kidger for Equinox and Google DeepMind for JMP, which served as the foundation for this implementation.

The authors acknowledge the computing time provided by the NHR Center NHR4CES at RWTH Aachen University (project number p0021919), funded by the Federal Ministry of Education and Research, and participating state governments through the GWK resolutions for national high performance computing at universities.

\begin{thebibliography}{9}
\bibitem{jmp} 
JMP: JAX Mixed Precision
\newblock \url{https://github.com/google-deepmind/jmp}

\bibitem{equinox}
Equinox: Neural Networks in JAX
\newblock \url{https://docs.kidger.site/equinox/}

\bibitem{nvidia_amp}
NVIDIA Automatic Mixed Precision
\newblock \url{https://developer.nvidia.com/automatic-mixed-precision}

\bibitem{mixed_precision_paper}
P. Micikevicius et al.
\newblock ``Mixed Precision Training''
\newblock arXiv:1710.03740, 2017
\end{thebibliography}

\end{document}

### 2025-05-20 mpx/cast.py

@@ -131,6 +131,38 @@ def cast_to_half_precision(x: PyTree) -> PyTree:
     return cast_tree(x, half_precision_datatype())
 
 
+def cast_function(func, dtype, return_dtype=None):
+    """
+    Casts the function to the specified data type.
+    """
+
+    if return_dtype is None:
+        return_dtype = dtype
+
+    def wrapper(*args, **kwargs):
+        args_cast = []
+        for arg in args:
+            args_cast.append(cast_tree(arg, dtype))
+        args_cast = tuple(args_cast)
+
+        kwargs_cast = {}
+        for key, value in kwargs.items():
+            kwargs_cast[key] = cast_tree(value, dtype)
+
+        results = func(*args_cast, **kwargs_cast)
+
+        if type(results) == tuple:
+            results_converted = []
+            for r in results:
+                results_converted.append(cast_tree(r, return_dtype))
+            return tuple(results_converted)
+        elif eqx.is_array(results):
+            return cast_tree(results, return_dtype)
+        return results
+    
+    return wrapper
+
+
 def force_full_precision(func, return_dtype=jnp.float16):
     """
     A decorator to enforce full precision (float32) for the inputs and outputs of a function.
@@ -150,34 +182,4 @@ def force_full_precision(func, return_dtype=jnp.float16):
         # will be cast to the specified `return_dtype` if it is an array.
     """
 
-    def wrapper(*args, **kwargs):
-        args_full_precision = []
-        for arg in args:
-            if eqx.is_array(arg):
-                args_full_precision.append(arg.astype(jnp.float32))
-            else:
-                args_full_precision.append(arg)
-        args_full_precision = tuple(args_full_precision)
-
-        kwargs_full_precision = {}
-        for key, value in kwargs.items():
-            if eqx.is_array(value):
-                kwargs_full_precision[key] = value.astype(jnp.float32)
-            else:
-                kwargs_full_precision[key] = value
-
-        results = func(*args_full_precision, **kwargs_full_precision)
-
-        if type(results) == tuple:
-            results_converted = []
-            for r in results:
-                if eqx.is_array(r):
-                    results_converted.append(r.astype(return_dtype))
-                else:
-                    results_converted.append(r)
-            return tuple(results_converted)
-        elif eqx.is_array(results):
-            return results.astype(return_dtype)
-        return results
-    
-    return wrapper
+    return cast_function(func, jnp.float32, return_dtype)

### 2025-05-20 tests/test_cast.py

@@ -13,6 +13,7 @@ from mpx.cast import (
     cast_to_full_precision,
     cast_to_half_precision,
     force_full_precision,
+    cast_function,
 )
 from mpx.dtypes import HALF_PRECISION_DATATYPE
 
@@ -220,6 +221,92 @@ class TestCastFunctions(unittest.TestCase):
         self.assertEqual(result['int32'].dtype, jnp.int32)
         self.assertEqual(result['bool'].dtype, jnp.bool_)
         self.assertEqual(result['nested']['int32'].dtype, jnp.int32)
+
+
+class TestCastFunction(unittest.TestCase):
+    def setUp(self):
+        self.array_float32 = jnp.array([1.0, 2.0, 3.0], dtype=jnp.float32)
+        self.array_float16 = jnp.array([1.0, 2.0, 3.0], dtype=jnp.float16)
+        self.array_bfloat16 = jnp.array([1.0, 2.0, 3.0], dtype=jnp.bfloat16)
+
+    def test_basic_function_casting(self):
+        def add(x, y):
+            return x + y
+
+        # Test with float32 inputs
+        casted_add = cast_function(add, jnp.float32)
+        result = casted_add(self.array_float32, self.array_float32)
+        self.assertEqual(result.dtype, jnp.float32)
+
+        # Test with float16 inputs
+        casted_add = cast_function(add, jnp.float16)
+        result = casted_add(self.array_float16, self.array_float16)
+        self.assertEqual(result.dtype, jnp.float16)
+
+    def test_tuple_return_values(self):
+        def operations(x, y):
+            return x + y, x * y
+
+        # Test with float32 inputs
+        casted_ops = cast_function(operations, jnp.float32)
+        result1, result2 = casted_ops(self.array_float32, self.array_float32)
+        self.assertEqual(result1.dtype, jnp.float32)
+        self.assertEqual(result2.dtype, jnp.float32)
+
+        # Test with float16 inputs
+        casted_ops = cast_function(operations, jnp.float16)
+        result1, result2 = casted_ops(self.array_float16, self.array_float16)
+        self.assertEqual(result1.dtype, jnp.float16)
+        self.assertEqual(result2.dtype, jnp.float16)
+
+    def test_different_input_output_dtypes(self):
+        def convert(x):
+            return x
+
+        casted_convert = cast_function(convert, jnp.float32, return_dtype=jnp.float16)
+        result = casted_convert(self.array_float16)
+        self.assertEqual(result.dtype, jnp.float16)
+
[…]

### 2025-05-20 doc/paper/main.tex
added pypi configs

@@
 \usepackage[utf8]{inputenc}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage{amsmath}
@@
 
-\title{MPX: Mixed Precision Training for JAX}
+\newcommand{\mpx}{\textsc{MPX}}
+
+\title{\mpx{}: Mixed Precision Training for JAX}
 \author{}
@@
 \section{Introduction}
-This paper presents \textbf{mpx}, a tool for training JAX models using mixed precision.
-The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{equinox}.
-By leveraging Equinox's flexibility, mpx provides a solution that works with any PyTree structure.
+This paper presents \mpx{}, a tool for training JAX models using mixed precision.
+The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{kidger2021equinox}.
+By leveraging Equinox's flexibility, \mpx{} provides a solution that works with any PyTree structure.
 
@@
 \section{Implementation Details}
-mpx is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
+\mpx{} is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
 For this, it provides several functions that allow to cast
 
-The mpx library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:
+The \mpx{} library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:
 
 \begin{enumerate}
-    \item \textbf{Transformations to Cast PyTrees}: mpx features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}.
+    \item \textbf{Transformations to Cast PyTrees}: \mpx{} features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}.
 All these functions cast all leaves of the input that are JAX arrays and of type float to the corresponding float datatype.
 All other leaves, including arrays that are of non-float types, like int32, remain unchanged.
-    \item \textbf{Transformations to Cast Functions}: mpx contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions.
+    \item \textbf{Transformations to Cast Functions}: \mpx{} contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions.
 This transformation returns a function that casts all its inputs to the desired input datatype (using \texttt{cast\_tree(tree, dtype)}), calls the function and then casts the outputs of the function.
-Moreover, mpx contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision.
+Moreover, \mpx{} contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision.
 This is important as some operations, such as sum, mean or softmax, are sensitive to overflows when calculated in float16.
-    \item \textbf{Transformations to Cast Gradients}: mpx contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above).
+    \item \textbf{Transformations to Cast Gradients}: \mpx{} contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above).
 Additional to calculating the gradient, the functions also perform the automatic adaption of the loss scaling value.
@@
 \subsection{Optimizer}
-mpx works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
+\mpx{} works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
 The  \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} function handles model updates based on gradient finiteness.
-This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{mpx.optimizer\_update}.
+This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{\mpx{}.optimizer\_update}.
 
 \section{Example}
-I t
+Here, we provide an example and show which parts in a training pipeline need to be changed for mixed precision training.
+
+\section{Model Implementation}
+For the largest part, the implementation of the model must not be changed. 
+As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\cite{flax2020github} or Equinox~\cite{kidger2021equinox} can be used in conjunction with \mpx{}.
 
@@
 
-\begin{thebibliography}{9}
-\bibitem{jmp} 
-JMP: JAX Mixed Precision
-\newblock \url{https://github.com/google-deepmind/jmp}
+\bibliographystyle{plain}
+\bibliography{references}
 
-\bibitem{equinox}
-Equinox: Neural Networks in JAX
-\newblock \url{https://docs.kidger.site/equinox/}
+% \begin{thebibliography}{9}
+% \bibitem{jmp} 
+% JMP: JAX Mixed Precision
+% \newblock \url{https://github.com/google-deepmind/jmp}
 
-\bibitem{nvidia_amp}
-NVIDIA Automatic Mixed Precision
-\newblock \url{https://developer.nvidia.com/automatic-mixed-precision}
+% \bibitem{equinox}
+% Equinox: Neural Networks in JAX
+% \newblock \url{https://docs.kidger.site/equinox/}
 
-\bibitem{mixed_precision_paper}
-P. Micikevicius et al.
-\newblock ``Mixed Precision Training''
-\newblock arXiv:1710.03740, 2017
-\end{thebibliography}
+% \bibitem{nvidia_amp}
+% NVIDIA Automatic Mixed Precision
+% \newblock \url{https://developer.nvidia.com/automatic-mixed-precision}
+
+% \bibitem{mixed_precision_paper}
+% P. Micikevicius et al.
+% \newblock ``Mixed Precision Training''
+% \newblock arXiv:1710.03740, 2017
+% \end{thebibliography}
 

### 2025-05-20 mpx/__init__.py

@@ -2,6 +2,8 @@
 Mixed Precision for JAX - A library for mixed precision training in JAX
 """
 
+__version__ = "0.1.2"
+
 from .cast import (
     cast_tree,
     cast_to_float32,

### 2025-05-20 mpx/loss_scaling.py

@@ -73,7 +73,29 @@ def scaled(func: callable, scaling: 'DynamicLossScaling', has_aux: bool = False)
 
 
 class DynamicLossScaling(eqx.Module):
-    """Basic structure taken from jmp."""
+    """
+    Implements dynamic loss scaling for mixed precision training in JAX.
+    The basic structure is taken from jmp.
+    This class automatically adjusts the loss scaling factor during training to prevent
+    numerical underflow/overflow when using reduced precision (e.g., float16). The scaling
+    factor is increased periodically if gradients are finite, and decreased if non-finite
+    gradients are detected, within specified bounds.
+    Attributes:
+        loss_scaling (jnp.ndarray): Current loss scaling factor.
+        min_loss_scaling (jnp.ndarray): Minimum allowed loss scaling factor.
+        counter (jnp.ndarray): Counter for tracking update periods.
+        factor (int): Multiplicative factor for adjusting loss scaling.
+        period (int): Number of steps between potential increases of loss scaling.
+    Methods:
+        scale(tree):
+            Scales all leaves of a pytree by the current loss scaling factor.
+        unscale(tree):
+            Unscales all leaves of a pytree by the inverse of the current loss scaling factor,
+            casting the result to float32.
+        adjust(grads_finite: jnp.ndarray) -> 'DynamicLossScaling':
+            Returns a new DynamicLossScaling instance with updated loss scaling and counter,
+            depending on whether the gradients are finite.
+    """
     loss_scaling: jnp.ndarray
     min_loss_scaling: jnp.ndarray
     counter: jnp.ndarray

### 2025-05-22 doc/paper/main.tex
wrote first draft of paper

@@
-\documentclass[11pt,a4paper]{article}
-\usepackage[left=2cm,right=2cm]{geometry}
+\documentclass[10pt, a4paper, logo, onecolumn, internal, copyright]{dsme}
+
+\usepackage[authoryear, sort&compress, round]{natbib}
+\bibliographystyle{abbrvnat}
 
 \usepackage[utf8]{inputenc}
-\usepackage[hidelinks]{hyperref}
+
 \usepackage{amsmath}
@@
 \usepackage{enumitem}
+\usepackage{mdframed}
+
+\usepackage{caption}
+\usepackage{subcaption}
+
+\usepackage{siunitx}
+
+\DeclareCaptionType[fileext=los,placement={!ht}]{example}
 
@@
 
-\title{\mpx{}: Mixed Precision Training for JAX}
-\author{}
+\title{\mpx{}: mixed precision Training for JAX}
+\correspondingauthor{}
+\reportnumber{} % Leave blank if n/a
+
+\author[1]{Alexander Gräfe}
+\author[1]{Sebastian Trimpe}
+
+\affil[1]{RWTH Aachen University}
 \date{}
+
+\begin{abstract}
+    MPX is a mixed-precision training toolbox for JAX that simplifies and accelerates the training of large-scale neural networks while preserving model accuracy. 
+    MPX seamlessly integrates with popular toolboxes such as Equinox and Flax, allowing users to convert full-precision pipelines to mixed-precision versions with minimal modifications. 
+    By casting both inputs and outputs to half precision, and introducing a dynamic loss-scaling mechanism, MPX alleviates issues like gradient underflow and overflow that commonly arise in half precision computations.
+    Its design inherits critical features from JAX's type-promotion behavior, ensuring that operations take place in the correct precision and allowing for selective enforcement of full precision where needed (e.g., sums, means, or softmax).
+    MPX further provides wrappers for automatic creation and management of mixed-precision gradients and optimizers, enabling straightforward integration into existing JAX training pipelines.
+    \mpx{}'s source code, documentation, and usage examples are available at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
+\end{abstract}
 
@@
 \section{Introduction}
-This paper presents \mpx{}, a tool for training JAX models using mixed precision.
-The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{kidger2021equinox}.
-By leveraging Equinox's flexibility, \mpx{} provides a solution that works with any PyTree structure.
 
-\section{Basics of Mixed Precision Training}
-This section summarizes the original Mixed Precision method from NVIDIA's Automatic Mixed Precision \cite{nvidia_amp} and the paper by Micikevicius et al. \cite{mixed_precision_paper}.
+In recent years, JAX~\cite{deepmind2020jax} has become an increasingly popular tool for training neural networks. 
+A key feature of JAX is its just-in-time compilation, which significantly accelerates code execution on hardware accelerators such as GPUs and TPUs. 
+In addition, JAX supports sharding mechanisms for multi-device and multi-node training, making it easy to scale up the training of large models across entire computing clusters.
 
-Mixed Precision training involves performing most computations in the forward and backward passes of a neural network using 16-bit floating-point numbers. This approach offers two main advantages:
+Many libraries have been developed to extend JAX's capabilities, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
+However, JAX has so far lacked a reliable, up-to-date library to enable mixed-precision training.
+The only existing option, JMP\citep{jmp}, is no longer maintained and cannot handle arbitrary JAX PyTrees, which are commonly used in modern neural-network toolboxes like Equinox~\citep{kidger2021equinox}.
+
+Mixed precision training is indispensable for efficient training of large models~\cite{mixed_precision_paper}. 
+By executing both the forward and backward passes of a neural network in half-precision floating-point format (requiring only \qty{16}{\bit}), mixed-precision techniques reduce memory requirements, particularly for intermediate activations, and speed up training.
+This speedup arises from lower memory transfer overhead and from specialized half-precision tensor cores in case the used hardware features them. 
+Crucially, these benefits typically come without compromising accuracy.
+
+This work presents \mpx{}, a mixed-precision training toolbox for JAX. 
+\mpx{} builds on the JMP implementation and leverages Equinox's flexibility to overcome limitations in handling arbitrary PyTrees.
+It further introduces gradient transformations and optimizer wrappers, enabling seamless conversion of a full-precision pipeline into mixed precision with only minor modifications.
+\mpx{}'s source code, documentation, and usage examples are available at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
+
+Before detailing the core functionality of \mpx{}, we first provide an overview of mixed-precision training. 
+We then illustrate how \mpx{} can be integrated into existing workflows by presenting sample code that demonstrates its main features.
+
+\section{Basics of mixed precision Training}
+This section summarizes the original mixed precision method from NVIDIA's Automatic mixed precision \citep{mixed_precision_paper} and the paper by Micikevicius et al. \citep{mixed_precision_paper}.
+
+mixed precision training involves performing most computations in the forward and backward passes of a neural network using 16-bit floating-point numbers. This approach offers two main advantages:
 \begin{itemize}
@@
 \subsection{Loss Scaling}
-A critical component of successful Mixed Precision training is loss scaling.
+A critical component of successful mixed precision training is loss scaling.
 Due to float16's decreased resolution, small gradients may be cast to zero, negatively impacting training performance.
@@
 \begin{itemize}
-    \item If scaled gradients exceed float16 range (inf), reduce scaling and skip model update
-    \item If scaled gradients remain within float16 range for an extended period, increase scaling
+    \item If scaled gradients exceed float16 range (inf), reduce scaling and skip model update.
+    \item If scaled gradients remain within float16 range for an extended period, increase scaling.
 \end{itemize}
@@
 \section{Implementation Details}
-\mpx{} is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
-For this, it provides several functions that allow to cast
 
-The \mpx{} library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:
+\begin{figure}[t]
+    \centering
+    \caption{Transformation of training pipelines via MPX.
+An existing training pipeline (Top figure) can be transformed, by applying \texttt{@mpx.filter\_grad} to the forward pass.
+It automatically casts all inputs to half precision and calculates the gradient via dynamic loss scaling.}
+    \label{fig:overview}
+    \begin{subfigure}[t]{0.99\textwidth}
+        \centering
+        \caption{Full Precision Training via Equinox}
+        \includegraphics[width=0.99\textwidth]{assets/EQXOverview.pdf}
+    \end{subfigure}\hspace{0.1cm}
+    \begin{subfigure}[t]{0.99\textwidth}
+        \centering
+        \caption{Half Precision Training via Equinox and \mpx{}}
+        \includegraphics[width=0.99\textwidth]{assets/MPXOverview.pdf}
+    \end{subfigure}\hspace{0.1cm}
+\end{figure}
+
+The \mpx{} library provides essential transformations for mixed precision training while retaining JAX's low-level flexibility. 
+Its design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are performed in the precision to which the inputs and outputs have been cast—provided that all constants in the function are positioned on the left side of the type-promotion lattice.
+As such, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision. 
+Concretely, it first defines functions to cast PyTrees (Section~\ref{sec:castingpytrees}), then uses these to cast both functions (Section~\ref{sec:castingfunctions}) and gradient calculations (Section~\ref{sec:castinggradients}) to mixed precision.
+In addition, we employ dynamic loss scaling~\citep{mixed_precision_paper} (Section~\ref{sec:scaling}) to stabilize mixed precision training.
+
+\subsection{Transformations to Cast PyTrees}
+\label{sec:castingpytrees}
+
+\mpx{} provides several functions for casting arbitrary PyTrees to a desired floating-point data type. These include
+\texttt{cast\_tree(tree, dtype)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, and \texttt{cast\_to\_float32(x)}.
+All of these functions inspect each leaf of the input PyTree, and any leaf that is a JAX array of a floating type is cast to the specified data type.
+Other leaves, such as integer arrays, remain unchanged.
+Excluding integer arrays is important to avoid accidental casting of random generator keys.
+
+\subsection{Transformations to Cast Functions}
+\label{sec:castingfunctions}
+
+For casting entire functions, \mpx{} provides the transformation \texttt{cast\_function(func, dtype, return\_dtype=None)}. 
+This transformation returns a new function that first applies the specified input data type to all inputs (via \texttt{cast\_tree(tree, dtype)}), then calls the original function, and finally casts the outputs to the specified (optional) return data type.
+Furthermore, \mpx{} includes \texttt{force\_full\_precision(func, return\_dtype)}, which ensures computations are carried out in full precision. 
+This feature is particularly important for operations that are prone to overflow in \texttt{float16}, such as \texttt{sum}, \texttt{mean}, or \texttt{softmax}.
+
+\subsection{Automatic Loss Scaling}
+\label{sec:scaling}
+
+To facilitate the calculation of gradients via loss scaling, \mpx{} offers the class \texttt{DynamicLossScaling}, which extends \texttt{jmp.DynamicLossScaling}.
+This class manages dynamic loss scaling through the following methods:
+
+\begin{itemize}
+    \item \texttt{scale(x)}: Multiplies all floating-point leaves of the PyTree \texttt{x} by the current scaling factor.
+    \item \texttt{unscale(x)}: Divides all floating-point leaves of \texttt{x} by the scaling factor and casts them to full precision.
+    \item \texttt{adjust(grads\_finite)}: Updates the scaling factor based on the stability of the gradients, following the techniques described by \cite{nvidia_amp}.
+\end{itemize}
+
+Because \texttt{DynamicLossScaling} inherits from \texttt{eqx.Module}, it is itself a PyTree. 
+Consequently, it can be used within \texttt{jit}-compiled functions and also replicated or sharded across multiple devices for distributed training.
+
+\subsection{Transformations to Cast Gradient Calculation}
+\label{sec:castinggradients}
+
+\mpx{} provides Equinox-compatible versions of gradient-calculation functions, namely \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)}.
+These functions compute gradients with mixed precision and loss scaling (as described in Section~\ref{sec:scaling}) and also automatically adjust the loss scaling value. 
+They serve as drop-in replacements for the corresponding Equinox functions, allowing users to reuse existing training pipelines with minimal changes (cf. Section~\ref{sec:eqxvsmpxmodelupdate}).
+
+To use these functions, the model's forward pass and loss computation should be combined into a single function, which is then passed to \texttt{mpx.filter\_value\_and\_grad} in place of \texttt{eqx.filter\_value\_and\_grad}.
+The transformed function then operates as follows (Figure~\ref{fig:overview}):
 
 \begin{enumerate}
-    \item \textbf{Transformations to Cast PyTrees}: \mpx{} features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}.
-All these functions cast all leaves of the input that are JAX arrays and of type float to the corresponding float datatype.
-All other leaves, including arrays that are of non-float types, like int32, remain unchanged.
-    \item \textbf{Transformations to Cast Functions}: \mpx{} contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions.
-This transformation returns a function that casts all its inputs to the desired input datatype (using \texttt{cast\_tree(tree, dtype)}), calls the function and then casts the outputs of the function.
-Moreover, \mpx{} contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision.
-This is important as some operations, such as sum, mean or softmax, are sensitive to overflows when calculated in float16.
-    \item \textbf{Transformations to Cast Gradients}: \mpx{} contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above).
-Additional to calculating the gradient, the functions also perform the automatic adaption of the loss scaling value.
-    These drop-in replacements allow users to reuse their existing Equinox training pipelines without major changes to their structure (cf. Section todo).
+    \item Cast all input arguments to half precision (\texttt{float16} or \texttt{bfloat16}).
+    \item Call the original function (model forward pass and loss).
+    \item Scale the function's output by the current dynamic scaling factor.
+    \item Apply \texttt{equinox.filter\_grad} to obtain scaled gradients.
+    \item Unscale the gradients by dividing them by the scaling factor.
+    \item Check whether the gradients are finite.
+    \item Update the loss scaling based on the gradient's finiteness.
+    \item Return a tuple containing:
+    \begin{itemize}
+        \item The updated \texttt{scaling} object.
+        \item A boolean indicating if the gradients are finite (for the optimizer step).
+        \item The computed gradients, whose format is a PyTree with the same format as the model parameters.
+        \item Auxiliary values (if \texttt{has\_aux=True}).
+    \end{itemize}
 \end{enumerate}
 
-\subsection{Automatic Loss Scaling Implementation}
-The \texttt{DynamicLossScaling} class manages dynamic loss scaling with methods:
-\begin{itemize}
-    \item \texttt{scale(x)}: Apply current loss scaling factor
-    \item \texttt{unscale(x)}: Remove loss scaling factor
-    \item \texttt{adjust(grads\_finite)}: Update scaling factor based on gradient stability
-\end{itemize}
+\subsection{Optimizer}
 
-\subsection{Optimizer}
-\mpx{} works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
-The  \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} function handles model updates based on gradient finiteness.
-This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{\mpx{}.optimizer\_update}.
+Because MPX produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}. 
+However, as discussed above, it may be necessary to skip optimizer updates altogether if the gradients become infinite due to loss scaling. 
+The function \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} implements this logic and updates the model only when gradients are finite.
+Consequently, instead of calling \texttt{optimizer.update} followed by \texttt{eqx.apply\_updates} (as in a typical Equinox training pipeline), one can simply call \texttt{mpx.optimizer\_update}.
 
@@
 
-\section{Model Implementation}
+\subsection{Model Implementation}
+
+\begin{example}[h]
+    \centering
+    \caption{Implementation of multi-headed self-attention via MPX.
+For the largest part, the model can be implemented in common toolboxes (here, Equinox~\cite{kidger2021equinox}).
+Only critical function that should be implemented in full precision, like softmax or layernorm (due to mean and std calculations), have to be transformed via \mpx{}.}
+    \label{ex:update}
+    \begin{subfigure}[t]{0.98\textwidth}
+        \centering
+        \caption{Full Precision via Equinox}
+        \vspace{-0.4cm}
+        \begin{mdframed}[
+            backgroundcolor=white!10!black,
+            leftmargin=0.5cm,hidealllines=true,%
+            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
+        ]
+            % \fontsize{8pt}{8pt}\selectfont
+            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+grads = eqx.filter_grad(loss)(model, batch)
+updates, optimizer_state = optimizer.update(
+    grads, optimizer_state, eqx.filter(model, eqx.is_array))
+model = eqx.apply_updates(model, updates)
+            \end{lstlisting}
+        \end{mdframed}
+    \end{subfigure}\hspace{0.1cm}
+    \begin{subfigure}[t]{0.98\textwidth}
+        \centering
+        \vspace{0.2cm}
+        \caption{mixed precision via MPX}
+        \vspace{-0.4cm}
+        \begin{mdframed}[
+            backgroundcolor=white!10!black,
+            leftmargin=0.5cm,hidealllines=true,%
+            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
+        ]
+            % \fontsize{8pt}{8pt}\selectfont
+            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+loss_scaling, grads_finite, grads = mpx.filter_grad(loss, loss_scaling)(
+    model, batch)
+model, optimizer_state = mpx.optimizer_update(
+    model, optimizer, optimizer_state, grads,grads_finite)
+            \end{lstlisting}
+        \end{mdframed}
+    \end{subfigure}
+
+\end{example}
+
 For the largest part, the implementation of the model must not be changed. 
-As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\cite{flax2020github} or Equinox~\cite{kidger2021equinox} can be used in conjunction with \mpx{}.
+As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Listing~\ref{ex:model}).
+Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.
+
+\subsection{Model Updates}
+\label{sec:eqxvsmpxmodelupdate}
+Model updates via \mpx{} are very similar to model updates via Equinox (Figure~\ref{ex:model}). Gradients can be calulated via
+
+\begin{example}[h]
+    \centering
+    \begin{minipage}{0.95\linewidth}
+            \caption{Implementation of model updates via Equinox and \mpx{}. \mpx{} is designed such that its model update routine closely resembles the one of Equinox.}
+            \label{ex:model}
+        \begin{mdframed}[
+    backgroundcolor=white!10!black,
+    leftmargin=0.5cm,hidealllines=true,%
+  innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
+]
+        % \fontsize{8pt}{8pt}\selectfont
+            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+class MultiHeadAttentionBlock(eqx.Module):
+    dense_qs: eqx.nn.Linear
+    dense_ks: eqx.nn.Linear
+    dense_vs: eqx.nn.Linear
+    dense_o: eqx.nn.Linear
+    num_heads: int
+    layer_norm: eqx.nn.LayerNorm
+
+    def __init__(self, feature_dim, num_heads, key):
+        self.num_heads = num_heads
+        key, subkey = jax.random.split(key)
+        self.dense_qs = eqx.nn.Linear(
+            feature_dim, feature_dim, key=subkey)
+        # same for dense_ks, dense_vs, dense_o
+
+        self.layer_norm = eqx.nn.LayerNorm(feature_dim)
+
+    def attention(q, k, v):
+        attention_scores = q @ k.T / jnp.sqrt(q.shape[-1])
+        <@\textcolor{white}{attention\_scores = mpx.force\_full\_precision(}@>
+            <@\textcolor{white}{jax.nn.softmax, attention\_scores.dtype)(attention\_scores, axis=-1)}@>
+        return attention_scores @ v
+
+    def __call__(self, inputs):
+        <@\textcolor{white}{inputs\_after\_layernorm = jax.vmap(mpx.force\_full\_precision(}@>
+            <@\textcolor{white}{self.layer\_norm, inputs.dtype))(inputs)}@>
+        qs = jax.vmap(self.dense_qs)(inputs_after_layernorm)
+        qs = es.jax_einshape("n(hf)->hnf", qs, h=self.num_heads)
+        # same for ks and vs...
+
+        outputs = jax.vmap(self.attention, in_axes=(0, 0, 0))(qs, ks, vs)
+        outputs = es.jax_einshape("hnf->n(hf)", outputs)
+        outputs = jax.vmap(self.dense_o)(outputs)
+        outputs += inputs
+
+        return outputs
+        \end{lstlisting}
+        \end{mdframed}
+    \end{minipage}
+\end{example}
 
@@
 
-The authors acknowledge the computing time provided by the NHR Center NHR4CES at RWTH Aachen University (project number p0021919), funded by the Federal Ministry of Education and Research, and participating state governments through the GWK resolutions for national high performance computing at universities.
+The authors gratefully acknowledge the computing time provided to them at the NHR Center NHR4CES at RWTH Aachen University (project number p0021919).
+This is funded by the Federal Ministry of Education and Research, and the state governments participating on the basis of the resolutions of the GWK for national high performance computing at universities (\url{www.nhr-verein.de/unsere-partner}).
 
-\bibliographystyle{plain}
 \bibliography{references}
@@
 % \bibitem{jmp} 
-% JMP: JAX Mixed Precision
+% JMP: JAX mixed precision
 % \newblock \url{https://github.com/google-deepmind/jmp}
@@
 % \bibitem{nvidia_amp}
-% NVIDIA Automatic Mixed Precision
-% \newblock \url{https://developer.nvidia.com/automatic-mixed-precision}
+% NVIDIA Automatic mixed precision
+% \newblock \url{https://developer.nvidia.com/automatic-mixed precision}
 
@@
 % P. Micikevicius et al.
-% \newblock ``Mixed Precision Training''
+% \newblock ``mixed precision Training''
 % \newblock arXiv:1710.03740, 2017

### 2025-05-22 mpx/loss_scaling.py

@@ -113,7 +113,7 @@ class DynamicLossScaling(eqx.Module):
             self.counter = counter
 
     def scale(self, tree):
-        return jax.tree_util.tree_map(lambda x: x * self.loss_scaling[0], tree)
+        return jax.tree_util.tree_map(lambda x: x * self.loss_scaling[0].astype(jnp.float16), tree)
 
     def unscale(self, tree):
         inv_loss_scaling = 1 / self.loss_scaling

### 2025-05-26 mpx/__init__.py
grad_tools now has function calculate_scaled_grad that is used inside the functions filter_x. This allows us to do unit-tests on the gradient calculation

@@ -1,8 +1,11 @@
 """
 Mixed Precision for JAX - A library for mixed precision training in JAX
+
+# build: python3 -m build --sdist
+# upload: twine upload --repository pypi dist/* 
 """
 
-__version__ = "0.1.2"
+__version__ = "0.1.3"
 
 from .cast import (
     cast_tree,

### 2025-05-26 mpx/grad_tools.py

@@ -127,6 +127,22 @@ def filter_grad(func, scaling: loss_scaling.DynamicLossScaling, has_aux=False, u
     return wrapper
 
 
+def calculate_scaled_grad(func, scaling: loss_scaling.DynamicLossScaling, has_aux=False, use_mixed_precision=True) -> PyTree: 
+    def wrapper(*args, **kwargs):
+        if use_mixed_precision:
+                args_cast = tuple([cast.cast_to_half_precision(x) for x in args])
+                kwargs_cast = {k: cast.cast_to_half_precision(v) for k, v in kwargs.items()}
+                func_scaled = loss_scaling.scaled(func, scaling, has_aux=has_aux)
+        else:
+            args_cast = args
+            kwargs_cast = kwargs
+            func_scaled = func
+
+        dfunc_scaled = eqx.filter_value_and_grad(func_scaled, has_aux=has_aux)
+        return dfunc_scaled(*args_cast, **kwargs_cast)
+    return wrapper
+
+
 def filter_value_and_grad(func, scaling: loss_scaling.DynamicLossScaling, has_aux=False, use_mixed_precision=True) -> PyTree:
     """
     Wraps a function to compute its value and gradient with support for mixed precision
@@ -155,19 +171,10 @@ def filter_value_and_grad(func, scaling: loss_scaling.DynamicLossScaling, has_au
     """
 
     def wrapper(*args, **kwargs):
-        if use_mixed_precision:
-            args_cast = tuple([cast.cast_to_half_precision(x) for x in args])
-            kwargs_cast = {k: cast.cast_to_half_precision(v) for k, v in kwargs.items()}
-            func_scaled = loss_scaling.scaled(func, scaling, has_aux=has_aux)
-        else:
-            args_cast = args
-            kwargs_cast = kwargs
-            func_scaled = func
-
-        dfunc_scaled = eqx.filter_value_and_grad(func_scaled, has_aux=has_aux)
+        dfunc_scaled = calculate_scaled_grad(func, scaling=scaling, has_aux=has_aux, use_mixed_precision=use_mixed_precision)
 
         if has_aux:
-            (value, aux), grad = dfunc_scaled(*args_cast, **kwargs_cast)
+            (value, aux), grad = dfunc_scaled(*args, **kwargs)
             if use_mixed_precision:
                 grads_finite = loss_scaling.all_finite(grad)
                 loss_scaling_new = scaling.adjust(grads_finite)
[…]

### 2025-05-26 tests/test_grad_tools.py

@@ -3,7 +3,7 @@ import jax
 import jax.numpy as jnp
 import equinox as eqx
 import optax
-from mpx.grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update
+from mpx.grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update, calculate_scaled_grad
 from mpx.loss_scaling import DynamicLossScaling
 
 
@@ -48,6 +48,47 @@ class TestGradTools(unittest.TestCase):
         result = select_tree(jnp.array(False), tree_a, tree_b)
         self.assertTrue(jnp.allclose(result['a'], tree_b['a']))
         self.assertTrue(jnp.allclose(result['b'], tree_b['b']))
+
+    def test_calculate_scaled_grad(self):
+        """Test the calculate_scaled_grad function"""
+
+        def loss_fn(model, x, y):
+            pred = model(x)
+            return jnp.mean((pred - y) ** 2)
+
+        x = jnp.array([1.0, 2.0])
+        y = jnp.array([2.0])
+
+        # Test without aux
+        scaled_grad_fn = calculate_scaled_grad(loss_fn, self.loss_scaling)
+        value, grad = scaled_grad_fn(self.model, x, y)
+        self.assertIsInstance(value, jnp.ndarray)
+        self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float16)
+        self.assertTrue(grad.bias.dtype == jnp.float16)
+
+        # Test with aux
+        def loss_fn_with_aux(model, x, y):
+            pred = model(x)
+            loss = jnp.mean((pred - y) ** 2)
+            return loss, {'pred': pred}
+
+        scaled_grad_fn = calculate_scaled_grad(loss_fn_with_aux, self.loss_scaling, has_aux=True)
+        (value, aux), grad = scaled_grad_fn(self.model, x, y)
+        self.assertIsInstance(value, jnp.ndarray)
+        self.assertIsInstance(aux, dict)
+        self.assertIn('pred', aux)
+        self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float16)
+        self.assertTrue(grad.bias.dtype == jnp.float16)
+
+        # Test with use_mixed_precision=False
+        scaled_grad_fn = calculate_scaled_grad(loss_fn, self.loss_scaling, use_mixed_precision=False)
[…]

### 2025-05-26 mpx/__init__.py
test_grad_tools now test for dtypes

@@ -5,7 +5,7 @@ Mixed Precision for JAX - A library for mixed precision training in JAX
 # upload: twine upload --repository pypi dist/* 
 """
 
-__version__ = "0.1.3"
+__version__ = "0.1.4"
 
 from .cast import (
     cast_tree,

### 2025-05-26 tests/test_grad_tools.py

@@ -105,6 +105,8 @@ class TestGradTools(unittest.TestCase):
         
         self.assertTrue(bool(grads_finite))
         self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float32)
+        self.assertTrue(grad.bias.dtype == jnp.float32)
         
         # Test with infinite gradients and no aux
         def bad_loss_fn(model, x, y):
@@ -114,6 +116,8 @@ class TestGradTools(unittest.TestCase):
         grad_fn = filter_grad(bad_loss_fn, self.loss_scaling)
         loss_scaling_new, grads_finite, grad = grad_fn(self.model, x, y)
         self.assertFalse(bool(grads_finite))
+        self.assertTrue(grad.weight.dtype == jnp.float32)
+        self.assertTrue(grad.bias.dtype == jnp.float32)
 
         # Test with finite gradients and aux
         # Test with aux
@@ -128,6 +132,8 @@ class TestGradTools(unittest.TestCase):
         self.assertIn('pred', aux)
         self.assertTrue(grads_finite)
         self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float32)
+        self.assertTrue(grad.bias.dtype == jnp.float32)
 
     def test_filter_value_and_grad(self):
         """Test the filter_value_and_grad function"""
@@ -145,6 +151,8 @@ class TestGradTools(unittest.TestCase):
         self.assertIsInstance(value, jnp.ndarray)
         self.assertTrue(grads_finite)
         self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float32)
+        self.assertTrue(grad.bias.dtype == jnp.float32)
         
         # Test with aux
         def loss_fn_with_aux(model, x, y):
@@ -160,6 +168,8 @@ class TestGradTools(unittest.TestCase):
         self.assertIn('pred', aux)
         self.assertTrue(grads_finite)
         self.assertIsInstance(grad, SimpleModel)
+        self.assertTrue(grad.weight.dtype == jnp.float32)
+        self.assertTrue(grad.bias.dtype == jnp.float32)
 
     def test_optimizer_update(self):
         """Test the optimizer_update function"""
@@ -181,6 +191,8 @@ class TestGradTools(unittest.TestCase):
         
         self.assertIsInstance(new_model, type(self.model))
[…]

### 2025-05-26 examples/train_vit.py
Cosmetic changes.
1. dtypes can now be set via strings.
2. Scaling is now a scalar and not an array with shape (1,)
3. Example now has a lower resolution,

@@ -1,6 +1,7 @@
 import itertools
 import os
-# os.environ['XLA_PYTHON_CLIENT_PREALLOCATE'] = 'false'
+# Disable XLA preallocation so one can see the memory consumed via nvidia-smi or nvitop (https://github.com/XuehaiPan/nvitop).
+os.environ['XLA_PYTHON_CLIENT_PREALLOCATE'] = 'false'
 
 import optax
 from tqdm import tqdm
@@ -188,9 +189,8 @@ def main(config):
         data = data.as_numpy_iterator()
         return data
 
-    # we make the resolution way too high for CIFAR100, but this is just for testing and to force the training to use a lot of memory.
-    train_dataset = init_tf_dataloader_image(train_data_source, config["batch_size"], config["num_epochs"], 0, 224)
-    val_dataset = init_tf_dataloader_image(val_data_source, config["batch_size"], config["num_epochs"], 0, 224)
+    train_dataset = init_tf_dataloader_image(train_data_source, config["batch_size"], config["num_epochs"], 0, 32)
+    val_dataset = init_tf_dataloader_image(val_data_source, config["batch_size"], config["num_epochs"], 0, 32)
 
     #########################################
     # Sharding
@@ -218,7 +218,7 @@ def main(config):
     model = filtered_device_put(model, replicated_sharding)
 
     if config["train_mixed_precision"]:
-        loss_scaling = mpx.DynamicLossScaling(loss_scaling=jnp.ones((1,), dtype=jnp.float32) * int((2 - 2**(-10)) * 2**15), min_loss_scaling=jnp.ones((1,), dtype=jnp.float32) * 1.0, period=2000)
+        loss_scaling = mpx.DynamicLossScaling(loss_scaling=mpx.FLOAT16_MAX, min_loss_scaling=jnp.ones((), dtype=jnp.float32) * 1.0, period=2000)
         loss_scaling = filtered_device_put(loss_scaling, replicated_sharding)
     else:
         loss_scaling = None
@@ -322,7 +322,7 @@ def main(config):
 
 if __name__ == "__main__":
     config = {
-        "train_mixed_precision": True,
+        "train_mixed_precision": False,
         "batch_size": 256,  
         "num_epochs": 10,
         "num_features": 256,

### 2025-05-26 examples/vit.py

@@ -42,7 +42,7 @@ class VIT(eqx.Module):
                  ):
         
         key, subkey = jax.random.split(key)
-        self.input_patch_length = 16
+        self.input_patch_length = 4
         self.num_channels = 3
         input_dim = self.num_channels * self.input_patch_length**2
         self.input_dim = input_dim

### 2025-05-26 mpx/__init__.py

@@ -16,7 +16,7 @@ from .cast import (
     cast_to_half_precision,
     force_full_precision,
 )
-from .dtypes import half_precision_datatype, set_half_precision_datatype
+from .dtypes import half_precision_datatype, set_half_precision_datatype, FLOAT16_MAX, BFLOAT16_MAX
 from .loss_scaling import DynamicLossScaling, all_finite, scaled
 from .grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update
 

### 2025-05-26 mpx/dtypes.py

@@ -1,6 +1,8 @@
 import jax.numpy as jnp
 
 HALF_PRECISION_DATATYPE = jnp.float16
+FLOAT16_MAX = jnp.ones([], dtype=jnp.float32) * (2 - 2**(-10)) * 2**15
+BFLOAT16_MAX = jnp.array([((2**8 - 1) * 2**(120))], dtype=jnp.float32)[0]
 
 def set_half_precision_datatype(datatype):
     """
@@ -10,7 +12,17 @@ def set_half_precision_datatype(datatype):
         datatype: The datatype to set as half precision (e.g., jnp.float16).
     """
     global HALF_PRECISION_DATATYPE
-    HALF_PRECISION_DATATYPE = datatype
+    if isinstance(datatype, str):
+        if datatype == 'float16':
+            datatype = jnp.float16
+        elif datatype == 'bfloat16':
+            datatype = jnp.bfloat16
+        else:
+            raise ValueError(f"Unsupported datatype: {datatype}. Use 'float16' or 'bfloat16'.")
+    elif datatype in (jnp.float16, jnp.bfloat16):
+        HALF_PRECISION_DATATYPE = datatype
+    else:
+        raise TypeError("Datatype must be a string or in (jnp.float16, jnp.bfloat16).")
 
 def half_precision_datatype():
     return HALF_PRECISION_DATATYPE

### 2025-05-26 mpx/loss_scaling.py

@@ -103,29 +103,30 @@ class DynamicLossScaling(eqx.Module):
     period: int
 
     def __init__(self, loss_scaling: jnp.ndarray, min_loss_scaling: jnp.ndarray, factor: int = 2, period: int = 2000, counter=None):
+        assert loss_scaling.ndim == 0, "Expected scalar loss scaling"
+        assert min_loss_scaling.ndim == 0, "Expected scalar minimum loss scaling"
         self.loss_scaling = loss_scaling
         self.min_loss_scaling = min_loss_scaling
         self.factor = factor
         self.period = period
         if counter is None:
-            self.counter = jnp.zeros((1,), dtype=jnp.int32)
+            self.counter = jnp.zeros((), dtype=jnp.int32)
         else:
             self.counter = counter
 
     def scale(self, tree):
-        return jax.tree_util.tree_map(lambda x: x * self.loss_scaling[0].astype(jnp.float16), tree)
+        return jax.tree_util.tree_map(lambda x: x * self.loss_scaling.astype(jnp.float16), tree)
 
     def unscale(self, tree):
         inv_loss_scaling = 1 / self.loss_scaling
         inv_loss_scaling = inv_loss_scaling.astype(jnp.float32)   # cast to float32, so the result is float32 (otherwise the whole scaling point would be senseless)
-        return jax.tree_util.tree_map(lambda x: x * inv_loss_scaling[0], tree)
+        return jax.tree_util.tree_map(lambda x: x * inv_loss_scaling, tree)
     
     def adjust(self, grads_finite: jnp.ndarray) -> 'DynamicLossScaling':
         """Returns the next state dependent on whether grads are finite."""
         assert grads_finite.ndim == 0, "Expected boolean scalar"
 
         first_finite = lambda a, b: jax.lax.select(jnp.isfinite(a).all(), a, b)
-
         loss_scaling = jax.lax.select(
             grads_finite,
 

### 2025-05-26 tests/test_grad_tools.py

@@ -24,8 +24,8 @@ class TestGradTools(unittest.TestCase):
     def setUp(self):
         # Create test data
         self.loss_scaling = DynamicLossScaling(
-            loss_scaling=jnp.array([2**10], dtype=jnp.float32),
-            min_loss_scaling=jnp.array([2**-14], dtype=jnp.float32),
+            loss_scaling=jnp.ones((), dtype=jnp.float32) ** 2**10,
+            min_loss_scaling=jnp.ones((), dtype=jnp.float32) * 2**-14,
             factor=2,
             period=2000
         )
@@ -224,8 +224,8 @@ class TestGradTools(unittest.TestCase):
         # Test with finite gradients and no aux
         # we make the scaling inf. If we set use_mixed_precision to False, the scaling should not be applied
         inf_loss_scaling = DynamicLossScaling(
-            loss_scaling=jnp.array([jnp.inf], dtype=jnp.float32),
-            min_loss_scaling=jnp.array([2**-14], dtype=jnp.float32),
+            loss_scaling=jnp.ones((), dtype=jnp.float32) * jnp.inf,
+            min_loss_scaling=jnp.ones((), dtype=jnp.float32) * 2**-14,
             factor=2,
             period=2000
         )
@@ -277,8 +277,8 @@ class TestGradTools(unittest.TestCase):
         # Test without aux
         # we make the scaling inf. If we set use_mixed_precision to False, the scaling should not be applied
         inf_loss_scaling = DynamicLossScaling(
-            loss_scaling=jnp.array([jnp.inf], dtype=jnp.float32),
-            min_loss_scaling=jnp.array([2**-14], dtype=jnp.float32),
+            loss_scaling=jnp.ones((), dtype=jnp.float32) * jnp.inf,
+            min_loss_scaling=jnp.ones((), dtype=jnp.float32) * 2**-14,
             factor=2,
             period=2000
         )

### 2025-05-26 tests/test_loss_scaling.py

@@ -7,8 +7,8 @@ from mpx.loss_scaling import DynamicLossScaling, all_finite, scaled
 class TestLossScaling(unittest.TestCase):
     def setUp(self):
         self.loss_scaling = DynamicLossScaling(
-            loss_scaling=jnp.array([2**14], dtype=jnp.float32),
-            min_loss_scaling=jnp.array([2**-14], dtype=jnp.float32),
+            loss_scaling=jnp.ones((), dtype=jnp.float32) * 2**14,
+            min_loss_scaling=jnp.ones((), dtype=jnp.float32) * 2**-14,
             factor=2,
             period=2000
         )
@@ -50,15 +50,15 @@ class TestLossScaling(unittest.TestCase):
             scaling = scaling.adjust(jnp.array(True))
         
         # After one period, scaling should have doubled
-        self.assertAlmostEqual(scaling.loss_scaling[0], 2**15)
-        self.assertEqual(scaling.counter[0], 0)  # Counter should reset
+        self.assertAlmostEqual(scaling.loss_scaling, 2**15)
+        self.assertEqual(scaling.counter, 0)  # Counter should reset
 
         for _ in range(2000):  # One full period
             scaling = scaling.adjust(jnp.array(True))
 
         # After two periods, scaling should not have doubled, but clipped to maximum of float16
-        self.assertAlmostEqual(scaling.loss_scaling[0], (2 - 2**(-10)) * 2**15)
-        self.assertEqual(scaling.counter[0], 0)  # Counter should reset
+        self.assertAlmostEqual(scaling.loss_scaling, (2 - 2**(-10)) * 2**15)
+        self.assertEqual(scaling.counter, 0)  # Counter should reset
 
     def test_adjust_infinite_grads(self):
         """Test loss scaling adjustment with infinite gradients"""
@@ -67,30 +67,30 @@ class TestLossScaling(unittest.TestCase):
         scaling = scaling.adjust(jnp.array(False))
         
         # Scaling should be halved
-        self.assertEqual(scaling.loss_scaling[0], 2**13)
-        self.assertEqual(scaling.counter[0], 0)  # Counter should reset
+        self.assertEqual(scaling.loss_scaling, 2**13)
+        self.assertEqual(scaling.counter, 0)  # Counter should reset
 
     def test_scaling_clipping(self):
         """Test that scaling is properly clipped"""
         # Test maximum clipping
[…]

### 2025-05-26 examples/train_vit.py
example now shows the time a step takes

@@ -1,8 +1,10 @@
 import itertools
 import os
+import time
 # Disable XLA preallocation so one can see the memory consumed via nvidia-smi or nvitop (https://github.com/XuehaiPan/nvitop).
 os.environ['XLA_PYTHON_CLIENT_PREALLOCATE'] = 'false'
 
+import numpy as np
 import optax
 from tqdm import tqdm
 
@@ -116,14 +118,15 @@ def train_epoch(model: eqx.Module,
                 show_progress: bool) -> tuple[eqx.Module, PyTree, PRNGKeyArray]:
     loss_value = 0
     num_datapoints = 0
-    
+    training_time = 0
     for batch in tqdm(itertools.islice(train_dataset, num_batches), disable=not show_progress and False):
         batch = jax.device_put(batch, batch_sharding)
 
         key, subkey = jax.random.split(key)
 
         loss_batch = 0
-        model, optimizer_state, loss_scaling, loss_batch = make_step(model=model, 
+        start_time = time.time()
+        model, optimizer_state, loss_scaling, loss_batch, = make_step(model=model, 
                                                     optimizer=optimizer, 
                                                     optimizer_state=optimizer_state,
                                                     batch=batch,
@@ -134,10 +137,12 @@ def train_epoch(model: eqx.Module,
                                                     key=subkey,
                                                     train_mixed_precicion=train_mixed_precicion
                                                     )
+        
+        training_time += time.time() - start_time
         loss_value += loss_batch.astype(jnp.float32) * len(batch["input"])
         num_datapoints += len(batch["input"])
     
-    return model, optimizer_state, loss_scaling, (loss_value) / (num_datapoints)
+    return model, optimizer_state, loss_scaling, (loss_value) / (num_datapoints), training_time
 
 
 def eval_epoch(model: eqx.Module,
@@ -189,8 +194,8 @@ def main(config):
         data = data.as_numpy_iterator()
         return data
 
-    train_dataset = init_tf_dataloader_image(train_data_source, config["batch_size"], config["num_epochs"], 0, 32)
-    val_dataset = init_tf_dataloader_image(val_data_source, config["batch_size"], config["num_epochs"], 0, 32)
+    train_dataset = init_tf_dataloader_image(train_data_source, config["batch_size"], config["num_epochs"], 0, config["resolution"])
+    val_dataset = init_tf_dataloader_image(val_data_source, config["batch_size"], config["num_epochs"], 0, config["resolution"])
 
[…]

### 2025-05-26 examples/vit.py

@@ -31,6 +31,7 @@ class VIT(eqx.Module):
 
 
     def __init__(self,
+                 input_patch_length: Int,
                  output_dim: Int,
                  num_features: Int,
                  num_heads: Int,
@@ -42,7 +43,7 @@ class VIT(eqx.Module):
                  ):
         
         key, subkey = jax.random.split(key)
-        self.input_patch_length = 4
+        self.input_patch_length = input_patch_length
         self.num_channels = 3
         input_dim = self.num_channels * self.input_patch_length**2
         self.input_dim = input_dim

### 2025-05-26 mpx/__init__.py

@@ -5,7 +5,7 @@ Mixed Precision for JAX - A library for mixed precision training in JAX
 # upload: twine upload --repository pypi dist/* 
 """
 
-__version__ = "0.1.4"
+__version__ = "0.1.5"
 
 from .cast import (
     cast_tree,

### 2025-05-28 doc/paper/main.tex
temp

@@
 
+\usepackage{microtype}
+
+\usepackage{tikz}
+\usepackage{pgfplots}
+\pgfplotsset{compat=1.18}
+
 \usepackage{caption}
 \usepackage{subcaption}
+\captionsetup[subfigure]{justification=centering}
 
@@
 \newcommand{\mpx}{\textsc{MPX}}
+\newcommand{\capt}[1]{\mdseries{\emph{#1}}}
 
@@
 \begin{document}
-
+\sloppy
 \maketitle
@@
     \centering
-    \caption{Transformation of training pipelines via MPX.
-An existing training pipeline (Top figure) can be transformed, by applying \texttt{@mpx.filter\_grad} to the forward pass.
-It automatically casts all inputs to half precision and calculates the gradient via dynamic loss scaling.}
-    \label{fig:overview}
     \begin{subfigure}[t]{0.99\textwidth}
         \centering
+        \includegraphics[width=0.99\textwidth]{assets/EQXOverview.pdf}
         \caption{Full Precision Training via Equinox}
-        \includegraphics[width=0.99\textwidth]{assets/EQXOverview.pdf}
+        \vspace{0.2cm}
     \end{subfigure}\hspace{0.1cm}
@@
         \centering
+        \includegraphics[width=0.99\textwidth]{assets/MPXOverview.pdf}
         \caption{Half Precision Training via Equinox and \mpx{}}
-        \includegraphics[width=0.99\textwidth]{assets/MPXOverview.pdf}
     \end{subfigure}\hspace{0.1cm}
+    \caption{Transformation of training pipelines via MPX.
+\capt{An existing training pipeline (Top figure) can be transformed, by applying \texttt{@mpx.filter\_grad} to the forward pass.
+It automatically casts all inputs to half precision and calculates the gradient via dynamic loss scaling and \texttt{eqx.filter\_grad}.}}
+    \label{fig:overview}
 \end{figure}
@@
     \centering
-    \caption{Implementation of multi-headed self-attention via MPX.
-For the largest part, the model can be implemented in common toolboxes (here, Equinox~\cite{kidger2021equinox}).
-Only critical function that should be implemented in full precision, like softmax or layernorm (due to mean and std calculations), have to be transformed via \mpx{}.}
-    \label{ex:update}
     \begin{subfigure}[t]{0.98\textwidth}
         \centering
-        \caption{Full Precision via Equinox}
-        \vspace{-0.4cm}
         \begin{mdframed}[
@@
         \end{mdframed}
+        \vspace{-0.5cm}
+        \caption{Full Precision via Equinox}
+        \vspace{0.5cm}
     \end{subfigure}\hspace{0.1cm}
@@
         \centering
-        \vspace{0.2cm}
-        \caption{mixed precision via MPX}
-        \vspace{-0.4cm}
         \begin{mdframed}[
@@
         \end{mdframed}
+        \vspace{-0.5cm}
+        \caption{mixed precision via MPX}
     \end{subfigure}
-
+    \caption{Implementation of model updates via Equinox and \mpx{}. \mpx{} is designed such that its model update routine closely resembles the one of Equinox.}
+    \label{ex:update}
 \end{example}
@@
     \begin{minipage}{0.95\linewidth}
-            \caption{Implementation of model updates via Equinox and \mpx{}. \mpx{} is designed such that its model update routine closely resembles the one of Equinox.}
-            \label{ex:model}
     \begin{mdframed}[
@@
     \end{minipage}
+    \caption{Implementation of multi-headed self-attention via MPX.
+\capt{For the largest part, the model can be implemented in common toolboxes (here, Equinox~\cite{kidger2021equinox}).
+Only critical function that should be implemented in full precision, like softmax or layernorm (due to mean and std calculations), have to be transformed via \mpx{}.}}
+    \label{ex:model}
 \end{example}
+
+\section{Evaluation}
+We evaluate \mpx{} on two different machines: A mid-level desktop PC (AMD Ryzen 9 5950X, \qty{82}{\giga\byte} RAM, Nvidia GeForce RTX4070 GPU with \qty{12}{\giga\byte} VRAM) and a node on a high-performance cluster (CLAIX-2023, Intel Xeon 8468 Sapphire CPUs, \qty{512}{\giga\byte} RAM, four NVIDIA H100 GPUs with \qty{96}{\giga\byte} VRAM each connected via NVLink).
+On paper, the desktop PC's GPU half precision computation speed is the same for half and full precision, while the H100 GPUs has double the speed for half than for full precision.
+
+We train small vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}.
+On both machines, the transformer has a feature size is 256, the residual blocks have one hidden layer with 800 neurons.
+The desktop PC trains the ViT on Cifar100 and the high-performance cluster on ImageNet1k, equally dividing each batch between the GPUs using a data-parallel approach.
+We train without gradient accumulation, which would be another way to optimize runtime and memory consumption.
+We vary the number of batches and measure the time the machines need to for each training step on the entire dataset, excluding the times used for data loading.
+
+We measure the RAM the program uses by turning off automatic preallocation and setting ``XLA\_PYTHON\_CLIENT\_PREALLOCATE'' to ``platform'', which minimizes the VRAM consumption of JAX\footnote{\url{https://docs.jax.dev/en/latest/gpu_memory_allocation.html}}
+For training speed, we turn both off.
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[t]{0.49\linewidth}
+    \begin{tikzpicture}
+        \begin{axis}[
+            width=0.8\textwidth,
+            height=7cm,
+            xlabel={Batch Size},
+            ylabel={VRAM (\qty{}{\giga\byte})},
+            legend pos=south east,
+            grid=major,
+            grid style={dashed,gray!30},
+            xmin=0,
+            ymin=0
+        ]
+        % First graph: Full Precision
+        \addplot[
+            color=blue,
+            mark=o,
+            thick
+        ] coordinates {
+            (128, 2.290)
+            (256, 4.724)
+            (512, 8.860)
+        };
+        \addlegendentry{Full Precision}
+
+        % Second graph: Mixed Precision
+        \addplot[
+            color=red,
+            mark=square*,
+            thick,
+            dashed
+        ] coordinates {
+            (128, 1.52)
+            (256, 2.418)
+            (512, 4.900)
+            (768, 7.054)
+            (1024, 9.222)
+        };
+        \addlegendentry{Mixed Precision}
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Desktop Computer (RTX4070)}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\linewidth}
+    \begin{tikzpicture}
+        \begin{axis}[
+            width=0.8\textwidth,
+            height=7cm,
+            xlabel={Batch Size},
+            ylabel={VRAM (\qty{}{\giga\byte})},
+            legend pos=south east,
+            grid=major,
+            grid style={dashed,gray!30},
+            xmin=0,
+            ymin=0
+        ]
+        % First graph: Full Precision
+        \addplot[
+            color=blue,
+            mark=o,
+            thick
+        ] coordinates {
+            (128, 2.290)
+            (256, 4.724)
+            (512, 8.860)
+        };
+        \addlegendentry{Full Precision}
+
+        % Second graph: Mixed Precision
+        \addplot[
+            color=red,
+            mark=square*,
+            thick,
+            dashed
+        ] coordinates {
+            (128, 1.52)
+            (256, 2.418)
+            (512, 4.900)
+            (768, 7.054)
+            (1024, 9.222)
+        };
+        \addlegendentry{Mixed Precision}
+        \end{axis}
+    \end{tikzpicture}
+    \caption{High-Performance Cluster ($4\times$ H100)}
+    \end{subfigure}
+    \caption{Comparison of GPU VRAM consumed for full precision and mixed precision as a function of the number of batches.}
+    \label{fig:memory}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[t]{0.49\linewidth}
+    \begin{tikzpicture}
+        \begin{axis}[
+            width=0.8\textwidth,
+            height=7cm,
+            xlabel={Batch Size},
+            ylabel={Training Time per Epoch (s)},
+            legend pos=south east,
+            grid=major,
+            grid style={dashed,gray!30},
+            xmin=0,
+            ymin=0,
+        ]
+        % First graph: Full Precision
+        \addplot[
+            color=blue,
+            mark=o,
+            thick
+        ] coordinates {
+            (128, 19.7)
+            (256, 18.8)
+            (512, 19.7)
+        };
+        \addlegendentry{Full Precision}
+
+        % Second graph: Mixed Precision
+        \addplot[
+            color=red,
+            mark=square*,
+            thick,
+            dashed
+        ] coordinates {
+            (128, 12.6)
+            (256, 12.0)
+            (512, 11.4)
+            (768, 11.6)
+            (1024, 11.5)
+        };
+        \addlegendentry{Mixed Precision}
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Desktop Computer (RTX4070)}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.49\linewidth}
+    \begin{tikzpicture}
+        \begin{axis}[
+            width=0.8\textwidth,
+            height=7cm,
+            xlabel={Batch Size},
+            ylabel={Training Time per Epoch (s)},
+            legend pos=south east,
+            grid=major,
+            grid style={dashed,gray!30},
+            xmin=0,
+            ymin=0,
+        ]
+        % First graph: Full Precision
+        \addplot[
+            color=blue,
+            mark=o,
+            thick
+        ] coordinates {
+            (128, 19.7)
+            (256, 18.8)
+            (512, 19.7)
+        };
+        \addlegendentry{Full Precision}
+
+        % Second graph: Mixed Precision
+        \addplot[
+            color=red,
+            mark=square*,
+            thick,
+            dashed
+        ] coordinates {
+            (128, 12.6)
+            (256, 12.0)
+            (512, 11.4)
+            (768, 11.6)
+            (1024, 11.5)
+        };
+        \addlegendentry{Mixed Precision}
+        \end{axis}
+    \end{tikzpicture}
+    \caption{High-Performance Cluster ($4\times$ H100)}
+    \end{subfigure}
+    \caption{Comparison of training step times for full precision and mixed precision as a function of the number of batches.}
+    \label{fig:timing}
+\end{figure}
+
+Mixed precision training via \mpx{} reduces both RAM and processing times.
+For the desktop PC, the RAM is reduced by factor $1.8\times$ and processing time by $1.7\times$.
+The speedup is caused by reduced memory loading times as the RTX4070 features no computing speedup for half precision.
+
+\section{Conclusion}
+
+\mpx{} offers an easy-to-use solution for mixed-precision training in JAX.
+It is able to handling arbitrary PyTrees and provides a seamless integration with popular libraries like Equinox and Flax. 
+Through dynamic loss scaling and precision casting, MPX ensures efficient memory usage and accelerates training without compromising model accuracy. 
+The evaluation demonstrates significant reductions in RAM usage and processing times, particularly on hardware that supports half-precision computation. 
+Overall, MPX presents a compelling advancement in optimizing large-scale neural network training pipelines while maintaining flexibility and ease of use.
 

### 2025-05-28 examples/train_vit.py

@@ -2,7 +2,8 @@ import itertools
 import os
 import time
 # Disable XLA preallocation so one can see the memory consumed via nvidia-smi or nvitop (https://github.com/XuehaiPan/nvitop).
-os.environ['XLA_PYTHON_CLIENT_PREALLOCATE'] = 'false'
+# os.environ['XLA_PYTHON_CLIENT_PREALLOCATE'] = 'false'
+# os.environ['XLA_PYTHON_CLIENT_ALLOCATOR'] = 'platform'
 
 import numpy as np
 import optax
@@ -67,6 +68,7 @@ def batched_loss_acc_wrapper(model, batch, batch_sharding, replicated_sharding, 
     return loss, acc
 
 
+
 @eqx.filter_jit
 def make_step(model: eqx.Module, 
             optimizer: any, 
@@ -77,25 +79,55 @@ def make_step(model: eqx.Module,
             loss_scaling: mpx.DynamicLossScaling,
             train_mixed_precicion: bool,
             weight_regularization: Float,
-            key: PRNGKeyArray
+            key: PRNGKeyArray,
+            num_grad_accumulation_steps: Int = 1
               ) -> tuple[eqx.Module, PyTree, Float, PRNGKeyArray]:
     batch = eqx.filter_shard(batch, batch_sharding)
     model = eqx.filter_shard(model, replicated_sharding)
     optimizer_state = eqx.filter_shard(optimizer_state, replicated_sharding)
     
-
-    if train_mixed_precicion:
-        (loss_value, _), loss_scaling, grads_finite, grads = mpx.filter_value_and_grad(batched_loss_acc_wrapper, scaling=loss_scaling, has_aux=True)(
-            model, batch, batch_sharding, replicated_sharding, key, weight_regularization)
-        model, optimizer_state = mpx.optimizer_update(model, optimizer, optimizer_state, grads,grads_finite)
+    if num_grad_accumulation_steps == 1:
+        if train_mixed_precicion:
+            (loss_value, _), loss_scaling, grads_finite, grads = mpx.filter_value_and_grad(batched_loss_acc_wrapper, scaling=loss_scaling, has_aux=True)(
+                model, batch, batch_sharding, replicated_sharding, key, weight_regularization)
+            model, optimizer_state = mpx.optimizer_update(model, optimizer, optimizer_state, grads,grads_finite)
+        else:
+            (loss_value, _), grads = eqx.filter_value_and_grad(batched_loss_acc_wrapper, has_aux=True)(
[…]

### 2025-05-29 doc/paper/main.tex
continued doc

@@
 
-In recent years, JAX~\cite{deepmind2020jax} has become an increasingly popular tool for training neural networks. 
-A key feature of JAX is its just-in-time compilation, which significantly accelerates code execution on hardware accelerators such as GPUs and TPUs. 
-In addition, JAX supports sharding mechanisms for multi-device and multi-node training, making it easy to scale up the training of large models across entire computing clusters.
+In recent years, JAX~\citep{deepmind2020jax} has emerged as a popular tool for training neural networks. 
+A standout feature of JAX is its just-in-time compilation, which significantly accelerates code execution on hardware accelerators such as GPUs and TPUs. 
+Additionally, JAX supports scalable multi-device and multi-node training across computing clusters through user-friendly and efficient sharding mechanisms.
 
-Many libraries have been developed to extend JAX's capabilities, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
-However, JAX has so far lacked a reliable, up-to-date library to enable mixed-precision training.
-The only existing option, JMP\citep{jmp}, is no longer maintained and cannot handle arbitrary JAX PyTrees, which are commonly used in modern neural-network toolboxes like Equinox~\citep{kidger2021equinox}.
+To extend JAX's capabilities, several libraries have been developed, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
+Despite these advancements, one critical area where JAX falls short is in supporting mixed-precision training, a method essential for efficiently handling large models~\cite{mixed_precision_paper}. 
+Although JMP~\citep{jmp} was previously available for this purpose, it is no longer maintained and cannot process arbitrary JAX PyTrees commonly used in modern neural-network toolboxes.
 
-Mixed precision training is indispensable for efficient training of large models~\cite{mixed_precision_paper}. 
-By executing both the forward and backward passes of a neural network in half-precision floating-point format (requiring only \qty{16}{\bit}), mixed-precision techniques reduce memory requirements, particularly for intermediate activations, and speed up training.
-This speedup arises from lower memory transfer overhead and from specialized half-precision tensor cores in case the used hardware features them. 
-Crucially, these benefits typically come without compromising accuracy.
+Mixed-precision training plays a pivotal role in optimizing the performance of large model training by executing both forward and backward passes in half-precision floating-point format (\qty{16}{\bit}).
+This technique significantly reduces memory requirements for intermediate activations and speeds up training by minimizing memory transfer overheads while utilizing specialized half-precision tensor cores (when supported by hardware), all without compromising accuracy~\cite{mixed_precision_paper}.
 
-This work presents \mpx{}, a mixed-precision training toolbox for JAX. 
-\mpx{} builds on the JMP implementation and leverages Equinox's flexibility to overcome limitations in handling arbitrary PyTrees.
-It further introduces gradient transformations and optimizer wrappers, enabling seamless conversion of a full-precision pipeline into mixed precision with only minor modifications.
-\mpx{}'s source code, documentation, and usage examples are available at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
+To facilitate the straightforward implementation of mixed-precision training in JAX, we introduce \mpx{}, a mixed-precision training toolbox tailored for JAX. 
+\mpx{} builds upon the JMP implementation and leverages Equinox's flexibility to overcome limitations associated with handling arbitrary PyTrees. 
+It further introduces gradient transformations and optimizer wrappers that enable seamless conversion from a full-precision pipeline to mixed precision with minimal modifications.
+The source code, documentation, and usage examples for \mpx{} are accessible at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
 
-Before detailing the core functionality of \mpx{}, we first provide an overview of mixed-precision training. 
-We then illustrate how \mpx{} can be integrated into existing workflows by presenting sample code that demonstrates its main features.
+Before explaining the core functionality of \mpx{}, we provide an overview of mixed-precision training principles. 
+Subsequently, we demonstrate how \mpx{} can be integrated into existing workflows through sample code illustrating its main features.
+Finally, we evaluate the performance of \mpx{} in the context of training vision transformers.
 
 \section{Basics of mixed precision Training}
-This section summarizes the original mixed precision method from NVIDIA's Automatic mixed precision \citep{mixed_precision_paper} and the paper by Micikevicius et al. \citep{mixed_precision_paper}.
-
-mixed precision training involves performing most computations in the forward and backward passes of a neural network using 16-bit floating-point numbers. This approach offers two main advantages:
+In this section, we summarize the foundational principles of mixed precision training as outlined in the original method~\citep{mixed_precision_paper}. 
+Mixed precision training involves executing most computations during the forward and backward passes of a neural network using 16-bit floating-point numbers. 
+This approach offers two primary advantages: 
 \begin{itemize} 
-    \item Reduced GPU memory usage (approximately 50\% compared to full precision)
-    \item Potential speedup through decreased memory access times and utilization of specialized half-precision tensor cores of TPUs or GPUs.
+    \item Reduced GPU Memory Usage: Mixed precision can decrease memory usage by approximately \SI{50}{\percent} compared to full precision. 
+    \item Potential Speedup: This technique can lead to faster execution due to reduced memory access times and the utilization of specialized half-precision tensor cores available in TPUs or GPUs. 
 \end{itemize}
@@
 \subsection{Loss Scaling}
-A critical component of successful mixed precision training is loss scaling.
-Due to float16's decreased resolution, small gradients may be cast to zero, negatively impacting training performance.
-The solution involves:
+One downside of using half-precision floating-point numbers is their decreased resolution. 
+Consequently, small gradients may be rounded to zero, which can negatively impact training performance.
+
+To mitigate these effects, loss scaling has been proposed. 
+The main idea involves multiplying the loss by a factor greater than one and calculating the gradient of this scaled loss. 
+Consequently, the gradients are also scaled by this factor, making them larger and allowing them to be represented more accurately by half-precision numbers.
+
+However, selecting a constant scaling factor poses challenges. 
+A small scaling factor might still result in gradients being rounded to zero, while an excessively large scaling factor can cause overflows where gradients become infinite. 
+Additionally, different stages of training may benefit from varying scaling factors depending on the training state.
+
+Dynamic loss scaling addresses this issue by adapting the scaling factor at runtime.
+Its objective is to maximize the scaling factor for best accuracy with half-precision numbers while avoiding overflow. 
+This is achieved through a simple heuristic that periodically increases the scaling factor and decreases it when gradients turn infinite.
+
+In summary, mixed precision training with dynamic loss scaling involves the following steps:
 \begin{enumerate}
-    \item Scaling the loss by a factor $> 1$
-    \item Calculating gradients with the scaled loss
-    \item Converting gradients to float32
-    \item Dividing by the scaling factor to obtain original gradients
+    \item Calculate the prediction of the model in half-precision.
+    \item Calculate the loss and scale it by a factor greater than 1.
+    \item Calculate gradients of the scaled loss with respect to the model parameters.
+    \item Converting gradients to float32.
+    \item Divide them by the scaling factor to retrieve original gradients.
+    \item Update the scaling factor:
+    \begin{enumerate}[label=(\alph*)]
+        \item If scaled gradients exceed float16's representable range (resulting in infinity), reduce the scaling and skip updating model parameters. 
+        \item If scaled gradients remain within float16 range for an extended period, increase scaling.
+    \end{enumerate}
+    \item Update the model parameters.
 \end{enumerate}
 
-The scaling factor can be automatically adjusted using a simple heuristic:
+The primary goal of \mpx{} is to offer functionalities that streamline the implementation of all these steps in mixed precision training. 
+In the following section, we provide an overview of these features.
+
+\section{Implementation Details}
+
+The \mpx{} library offers essential transformations for mixed precision training while preserving JAX's low-level flexibility. 
+Similar to JMP~\cite{jmp}, implementationts design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice.
+Consequently, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision.
+
+Specifically, \mpx{} first defines functions for casting PyTrees (Section\ref{sec:castingpytrees}), then utilizes these functions to cast both general functions (Section\ref{sec:castingfunctions}) and gradient calculations (Section\ref{sec:castinggradients}) into mixed precision.
+The gradient calculations include dynamic loss scaling (Section~\ref{sec:scaling}) to stabilize mixed precision training.
+Moreover, \mpx{} provides wrappers around optimizers to exclude update steps where gradients are infinite (Section~\ref{sec:optimizer}).
+
+\subsection{Transformations to Cast PyTrees}
+\label{sec:castingpytrees}
+
+\mpx{} provides several functions for casting arbitrary PyTrees to a desired floating-point data type. These functions include:
+\texttt{cast\_tree(tree, dtype)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, and \texttt{cast\_to\_float32(x)}. 
+Each function inspects the leaves of the input PyTree, casting any leaf that is a JAX array with a floating-point type to the desired data type. 
+Other leaves, such as integer arrays, remain unchanged. 
+Excluding integer arrays is crucial to prevent accidental casting of random generator keys.
+
+\subsection{Transformations to Cast Functions}
+\label{sec:castingfunctions}
+
+For casting entire functions, \mpx{} provides the transformation \texttt{cast\_function(func, dtype, return\_dtype=None)}. 
+This transformation returns a new function that first applies the specified input data type to all inputs (via \texttt{cast\_tree(tree, dtype)}), then calls the original function, and finally casts the outputs to the specified (optional) return data type.
+Furthermore, \mpx{} includes \texttt{force\_full\_precision(func, return\_dtype)}, which ensures computations are carried out in full precision. 
+This feature is particularly important for operations that are prone to overflow in \texttt{float16}, such as \texttt{sum}, \texttt{mean}, or \texttt{softmax}.
+
+\subsection{Automatic Loss Scaling}
+\label{sec:scaling}
+
+To facilitate the calculation of gradients via loss scaling, \mpx{} offers the class \texttt{DynamicLossScaling}, which extends \texttt{jmp.DynamicLossScaling}. 
+This class manages dynamic loss scaling through the following methods:
+
 \begin{itemize}
-    \item If scaled gradients exceed float16 range (inf), reduce scaling and skip model update.
-    \item If scaled gradients remain within float16 range for an extended period, increase scaling.
+    \item \texttt{scale(x)}: Multiplies all floating-point leaves of the PyTree \texttt{x} by the current scaling factor.
+    \item \texttt{unscale(x)}: Divides all floating-point leaves of \texttt{x} by the scaling factor and casts them to full precision.
+    \item \texttt{adjust(grads\_finite)}: Updates the scaling factor based on the stability of the gradients, following the techniques described by~\cite{mixed_precision_paper}.
 \end{itemize}
 
-\section{Implementation Details}
+Because \texttt{DynamicLossScaling} inherits from \texttt{eqx.Module}, it is itself a PyTree. 
+Consequently, it can be used within \texttt{jit}-compiled functions and also be used for multi-device and multi-node training using replicated sharding.
 
@@
 
-The \mpx{} library provides essential transformations for mixed precision training while retaining JAX's low-level flexibility. 
-Its design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are performed in the precision to which the inputs and outputs have been cast—provided that all constants in the function are positioned on the left side of the type-promotion lattice.
-As such, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision. 
-Concretely, it first defines functions to cast PyTrees (Section~\ref{sec:castingpytrees}), then uses these to cast both functions (Section~\ref{sec:castingfunctions}) and gradient calculations (Section~\ref{sec:castinggradients}) to mixed precision.
-In addition, we employ dynamic loss scaling~\citep{mixed_precision_paper} (Section~\ref{sec:scaling}) to stabilize mixed precision training.
-
-\subsection{Transformations to Cast PyTrees}
-\label{sec:castingpytrees}
-
-\mpx{} provides several functions for casting arbitrary PyTrees to a desired floating-point data type. These include
-\texttt{cast\_tree(tree, dtype)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, and \texttt{cast\_to\_float32(x)}.
-All of these functions inspect each leaf of the input PyTree, and any leaf that is a JAX array of a floating type is cast to the specified data type.
-Other leaves, such as integer arrays, remain unchanged.
-Excluding integer arrays is important to avoid accidental casting of random generator keys.
-
-\subsection{Transformations to Cast Functions}
-\label{sec:castingfunctions}
-
-For casting entire functions, \mpx{} provides the transformation \texttt{cast\_function(func, dtype, return\_dtype=None)}. 
-This transformation returns a new function that first applies the specified input data type to all inputs (via \texttt{cast\_tree(tree, dtype)}), then calls the original function, and finally casts the outputs to the specified (optional) return data type.
-Furthermore, \mpx{} includes \texttt{force\_full\_precision(func, return\_dtype)}, which ensures computations are carried out in full precision. 
-This feature is particularly important for operations that are prone to overflow in \texttt{float16}, such as \texttt{sum}, \texttt{mean}, or \texttt{softmax}.
-
-\subsection{Automatic Loss Scaling}
-\label{sec:scaling}
-
-To facilitate the calculation of gradients via loss scaling, \mpx{} offers the class \texttt{DynamicLossScaling}, which extends \texttt{jmp.DynamicLossScaling}.
-This class manages dynamic loss scaling through the following methods:
-
-\begin{itemize}
-    \item \texttt{scale(x)}: Multiplies all floating-point leaves of the PyTree \texttt{x} by the current scaling factor.
-    \item \texttt{unscale(x)}: Divides all floating-point leaves of \texttt{x} by the scaling factor and casts them to full precision.
-    \item \texttt{adjust(grads\_finite)}: Updates the scaling factor based on the stability of the gradients, following the techniques described by \cite{nvidia_amp}.
-\end{itemize}
-
-Because \texttt{DynamicLossScaling} inherits from \texttt{eqx.Module}, it is itself a PyTree. 
-Consequently, it can be used within \texttt{jit}-compiled functions and also replicated or sharded across multiple devices for distributed training.
-
 \subsection{Transformations to Cast Gradient Calculation}
@@
 
-\mpx{} provides Equinox-compatible versions of gradient-calculation functions, namely \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)}.
-These functions compute gradients with mixed precision and loss scaling (as described in Section~\ref{sec:scaling}) and also automatically adjust the loss scaling value. 
-They serve as drop-in replacements for the corresponding Equinox functions, allowing users to reuse existing training pipelines with minimal changes (cf. Section~\ref{sec:eqxvsmpxmodelupdate}).
+\mpx{} provides Equinox-compatible versions of gradient-calculation functions: \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)}.
+These functions compute gradients using mixed precision and loss scaling (as described in Section\ref{sec:scaling}) while automatically adjusting the loss scaling value. 
+They act as drop-in replacements for the corresponding Equinox functions, enabling users to reuse existing training pipelines with minimal modifications (cf. Section\ref{sec:eqxvsmpxmodelupdate}).
 
@@
     \item Cast all input arguments to half precision (\texttt{float16} or \texttt{bfloat16}).
-    \item Call the original function (model forward pass and loss).
+    \item Execute the original function (model forward pass and loss).
     \item Scale the function's output by the current dynamic scaling factor.
@@
     \item Check whether the gradients are finite.
-    \item Update the loss scaling based on the gradient's finiteness.
+    \item Update the loss scaling based on the finiteness of gradients.
     \item Return a tuple containing:
@@
 \subsection{Optimizer}
+\label{sec:optimizer}
 
-Because MPX produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}. 
-However, as discussed above, it may be necessary to skip optimizer updates altogether if the gradients become infinite due to loss scaling. 
-The function \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} implements this logic and updates the model only when gradients are finite.
-Consequently, instead of calling \texttt{optimizer.update} followed by \texttt{eqx.apply\_updates} (as in a typical Equinox training pipeline), one can simply call \texttt{mpx.optimizer\_update}.
+Since \mpx{} produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}.
+However, as previously discussed, it may be necessary to skip optimizer updates if the gradients become infinite due to loss scaling. 
+The function \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} implements this logic by updating the model only when gradients are finite.
+
+Thus, instead of calling \texttt{optimizer.update} followed by \texttt{eqx.apply\_updates}, as in a typical Equinox training pipeline, one can simply use \texttt{mpx.optimizer\_update}.
 
 \section{Example}
-Here, we provide an example and show which parts in a training pipeline need to be changed for mixed precision training.
+In this section, we provide an example to demonstrate the changes required in a training pipeline for mixed precision training via \mpx{}.
 
 \subsection{Model Implementation}
-
-\begin{example}[h]
-    \centering
-    \begin{subfigure}[t]{0.98\textwidth}
-        \centering
-        \begin{mdframed}[
-            backgroundcolor=white!10!black,
-            leftmargin=0.5cm,hidealllines=true,%
-            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
-        ]
-            % \fontsize{8pt}{8pt}\selectfont
-            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
-grads = eqx.filter_grad(loss)(model, batch)
-updates, optimizer_state = optimizer.update(
-    grads, optimizer_state, eqx.filter(model, eqx.is_array))
-model = eqx.apply_updates(model, updates)
-            \end{lstlisting}
-        \end{mdframed}
-        \vspace{-0.5cm}
-        \caption{Full Precision via Equinox}
-        \vspace{0.5cm}
-    \end{subfigure}\hspace{0.1cm}
-    \begin{subfigure}[t]{0.98\textwidth}
-        \centering
-        \begin{mdframed}[
-            backgroundcolor=white!10!black,
-            leftmargin=0.5cm,hidealllines=true,%
-            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
-        ]
-            % \fontsize{8pt}{8pt}\selectfont
-            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
-loss_scaling, grads_finite, grads = mpx.filter_grad(loss, loss_scaling)(
-    model, batch)
-model, optimizer_state = mpx.optimizer_update(
-    model, optimizer, optimizer_state, grads,grads_finite)
-            \end{lstlisting}
-        \end{mdframed}
-        \vspace{-0.5cm}
-        \caption{mixed precision via MPX}
-    \end{subfigure}
-    \caption{Implementation of model updates via Equinox and \mpx{}. \mpx{} is designed such that its model update routine closely resembles the one of Equinox.}
-    \label{ex:update}
-\end{example}
-
-For the largest part, the implementation of the model must not be changed. 
-As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Listing~\ref{ex:model}).
-Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.
-
-\subsection{Model Updates}
-\label{sec:eqxvsmpxmodelupdate}
-Model updates via \mpx{} are very similar to model updates via Equinox (Figure~\ref{ex:model}). Gradients can be calulated via
 
@@
 
+For the largest part, the implementation of the model must not be changed.
+As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Example~\ref{ex:model}).
+Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.
+
+\subsection{Model Updates}
+\label{sec:eqxvsmpxmodelupdate}
+Model updates via \mpx{} closely resemble those performed with Equinox (Example~\ref{ex:update}). 
+Instead of calling \texttt{eqx.filter\_grad}, \texttt{mpx.filter\_grad} is used. Similarly, \texttt{mpx.optimizer\_update} replaces invoking the Optax optimizer followed by a model update.
+The remainder of the training step—such as loading batches, sharding the model and batch—remains unchanged. 
+All these steps can be compiled using tools like \texttt{eqx.filter\_jit}.
+
+\begin{example}[h]
+    \centering
+    \begin{subfigure}[t]{0.98\textwidth}
+        \centering
+        \begin{mdframed}[
+            backgroundcolor=white!10!black,
+            leftmargin=0.5cm,hidealllines=true,%
+            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
+        ]
+            % \fontsize{8pt}{8pt}\selectfont
+            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+grads = eqx.filter_grad(loss)(model, batch)
+updates, optimizer_state = optimizer.update(
+    grads, optimizer_state, eqx.filter(model, eqx.is_array))
+model = eqx.apply_updates(model, updates)
+            \end{lstlisting}
+        \end{mdframed}
+        \vspace{-0.5cm}
+        \caption{Full Precision via Equinox}
+        \vspace{0.5cm}
+    \end{subfigure}\hspace{0.1cm}
+    \begin{subfigure}[t]{0.98\textwidth}
+        \centering
+        \begin{mdframed}[
+            backgroundcolor=white!10!black,
+            leftmargin=0.5cm,hidealllines=true,%
+            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
+        ]
+            % \fontsize{8pt}{8pt}\selectfont
+            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+loss_scaling, grads_finite, grads = mpx.filter_grad(loss, loss_scaling)(
+    model, batch)
+model, optimizer_state = mpx.optimizer_update(
+    model, optimizer, optimizer_state, grads,grads_finite)
+            \end{lstlisting}
+        \end{mdframed}
+        \vspace{-0.5cm}
+        \caption{mixed precision via MPX}
+    \end{subfigure}
+    \caption{Implementation of model updates via Equinox and \mpx{}. \mpx{} is designed such that its model update routine closely resembles the one of Equinox.}
+    \label{ex:update}
+\end{example}
+
 \section{Evaluation}
-We evaluate \mpx{} on two different machines: A mid-level desktop PC (AMD Ryzen 9 5950X, \qty{82}{\giga\byte} RAM, Nvidia GeForce RTX4070 GPU with \qty{12}{\giga\byte} VRAM) and a node on a high-performance cluster (CLAIX-2023, Intel Xeon 8468 Sapphire CPUs, \qty{512}{\giga\byte} RAM, four NVIDIA H100 GPUs with \qty{96}{\giga\byte} VRAM each connected via NVLink).
-On paper, the desktop PC's GPU half precision computation speed is the same for half and full precision, while the H100 GPUs has double the speed for half than for full precision.
+We evaluate \mpx{} on two distinct machines: a mid-level desktop PC equipped with an AMD Ryzen 9 5950X, \qty{82}{\giga\byte} RAM, and an Nvidia GeForce RTX4070 GPU featuring \qty{12}{\giga\byte} VRAM; and a node from a high-performance cluster (CLAIX-2023) comprising Intel Xeon 8468 Sapphire CPUs, \qty{512}{\giga\byte} RAM, and four NVIDIA H100 GPUs connected via NVLink each with \qty{96}{\giga\byte} VRAM.
+On paper, the desktop PC's GPU maintains equivalent computation speeds for both half and full precision, whereas the H100 GPUs deliver double the speed for half precision compared to full precision.
 
-We train small vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}.
-On both machines, the transformer has a feature size is 256, the residual blocks have one hidden layer with 800 neurons.
-The desktop PC trains the ViT on Cifar100 and the high-performance cluster on ImageNet1k, equally dividing each batch between the GPUs using a data-parallel approach.
-We train without gradient accumulation, which would be another way to optimize runtime and memory consumption.
-We vary the number of batches and measure the time the machines need to for each training step on the entire dataset, excluding the times used for data loading.
+We train vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}.
+On the desktop PC, the transformer features a size of 256 with residual blocks containing one hidden layer of 800 neurons.
+T
+The cluster's transformer mirrors ViT-Base dimensions, with a feature size of 768 and residual blocks having one hidden layer of 3072 neurons.
 
-We measure the RAM the program uses by turning off automatic preallocation and setting ``XLA\_PYTHON\_CLIENT\_PREALLOCATE'' to ``platform'', which minimizes the VRAM consumption of JAX\footnote{\url{https://docs.jax.dev/en/latest/gpu_memory_allocation.html}}
-For training speed, we turn both off.
+The desktop PC trains the ViT on CIFAR100 while the high-performance cluster handles ImageNet1k, dividing each batch equally across GPUs using a data-parallel approach.
+We train without gradient accumulation, which could further optimize runtime and memory consumption
+
+We vary the number of batches and measure the time required for each training step across the entire dataset, excluding data loading times. 
+RAM usage is measured by disabling automatic preallocation and setting "XLA\_PYTHON\_CLIENT\_PREALLOCATE" to "platform," minimizing JAX's VRAM consumption\footnote{\url{https://docs.jax.dev/en/latest/gpu_memory_allocation.html}}.
+For training speed assessment, both settings are turned off.
 
@@
         ] coordinates {
-            (128, 19.7)
-            (256, 18.8)
-            (512, 19.7)
+            (512, 595.5616071701049)
+            (1024, 452.3)
+            (2048, 428.24)
+            % (4096, 153.62)
+            % (5120, 152)
+            % (6144, 149.53)
+            % (7168, 153.88)
         };
@@
         ] coordinates {
-            (128, 12.6)
-            (256, 12.0)
-            (512, 11.4)
-            (768, 11.6)
-            (1024, 11.5)
+            (512, 435.09)
+            (1024, 325.61)
+            (2048, 272.48)
+            % (4096, 117.6)
+            % (5120, 113.62)
+            % (6144, 109.06)
+            % (7168, 104.51)
+            % (8192, 102.68)
         };
@@
 For the desktop PC, the RAM is reduced by factor $1.8\times$ and processing time by $1.7\times$.
-The speedup is caused by reduced memory loading times as the RTX4070 features no computing speedup for half precision.
+The speedup is caused by reduced memory loading times as the RTX4070 streaming-multiprocessors feature no computing speedup for half precision.
+The cluster features a similar behavior, where both RAM and processing times are reduced.
+RAM is reduced by factor $1.8\times$ and processing time by up to $1.57\times$.
+
+These experiments demonstrate that mixed precision via \mpx{} allows to significantly speed up training.
+Moreover, it significantly reduces VRAM usage.
 

### 2025-05-30 mpx/grad_tools.py
fixed a bug, where mpc crashed, when a leaf of a pytree was none

@@ -68,7 +68,7 @@ def select_tree(pred: jnp.ndarray, a: PyTree, b: PyTree) -> PyTree:
         else:
             return x1
 
-    return jax.tree_util.tree_map(_select_leaf, a, b)
+    return jax.tree_util.tree_map(lambda x, y: None if x is None else _select_leaf(x, y), a, b, is_leaf=lambda x: x is None)
 
 
 def filter_grad(func, scaling: loss_scaling.DynamicLossScaling, has_aux=False, use_mixed_precision=True) -> PyTree:

### 2025-05-30 tests/test_grad_tools.py

@@ -11,10 +11,12 @@ from mpx.loss_scaling import DynamicLossScaling
 class SimpleModel(eqx.Module):
     weight: jnp.ndarray
     bias: jnp.ndarray
+    none_leaf: None
     
     def __init__(self):
         self.weight = jnp.array([1.0, 2.0], dtype=jnp.float32)
         self.bias = jnp.array([0.5], dtype=jnp.float32)
+        self.none_leaf = None  # Example of a non-array leaf (lead to errors in the past)
     
     def __call__(self, x):
         return jnp.dot(x, self.weight) + self.bias

### 2025-05-30 mpx/grad_tools.py
reverted last change

@@ -68,7 +68,7 @@ def select_tree(pred: jnp.ndarray, a: PyTree, b: PyTree) -> PyTree:
         else:
             return x1
 
-    return jax.tree_util.tree_map(lambda x, y: None if x is None else _select_leaf(x, y), a, b, is_leaf=lambda x: x is None)
+    return jax.tree_util.tree_map(_select_leaf, a, b)
 
 
 def filter_grad(func, scaling: loss_scaling.DynamicLossScaling, has_aux=False, use_mixed_precision=True) -> PyTree:

### 2025-06-10 doc/paper/main.tex
temp

@@
-\documentclass[10pt, a4paper, logo, onecolumn, internal, copyright]{dsme}
+\documentclass[copyright, logo]{dsme}
 
@@
 
-\usepackage[utf8]{inputenc}
+%% Title
+\title{\mpx{}: Mixed Precision Training for JAX}
 
+\usepackage{authblk}
+
+\author[1]{Alexander Gräfe}
+\author[1]{Sebastian Trimpe}
+
+\renewcommand{\shortauthors}{Gräfe et al.}
+
+\affil[1]{Institute for Data Science in Mechanical Engineering (DSME), RWTH Aachen University}
+
+\codelabel{\faGithub} % defaults to "Code:"
+\codeurl{https://github.com/Data-Science-in-Mechanical-Engineering/mixed_precision_for_JAX}
+\codeurlshort{mixed\_precision\_for\_JAX}
+
+% \date{}
+
+% user packages
 \usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{listings}
-\usepackage{xcolor}
 \usepackage{enumitem}
 \usepackage{mdframed}
-
-\usepackage{microtype}
 
@@
 
-\title{\mpx{}: mixed precision Training for JAX}
-\correspondingauthor{}
-\reportnumber{} % Leave blank if n/a
-
-\author[1]{Alexander Gräfe}
-\author[1]{Sebastian Trimpe}
-
-\affil[1]{RWTH Aachen University}
-\date{}
-
+\begin{document}
+\sloppy
+\maketitle
 \begin{abstract}
-    MPX is a mixed-precision training toolbox for JAX that simplifies and accelerates the training of large-scale neural networks while preserving model accuracy. 
+Mixed-precision training has emerged as an indispensable tool for enhancing the efficiency of neural network training in recent years. 
+    Concurrently, JAX has grown in popularity as a versatile machine learning toolbox.
+    However, it currently lacks robust support for mixed-precision training.
+    We propose MPX, a mixed-precision training toolbox for JAX that simplifies and accelerates the training of large-scale neural networks while preserving model accuracy. 
     MPX seamlessly integrates with popular toolboxes such as Equinox and Flax, allowing users to convert full-precision pipelines to mixed-precision versions with minimal modifications. 
@@
     MPX further provides wrappers for automatic creation and management of mixed-precision gradients and optimizers, enabling straightforward integration into existing JAX training pipelines.
-    \mpx{}'s source code, documentation, and usage examples are available at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
+    \mpx{}'s source code, documentation, and usage examples are available at \url{github.com/Data-Science-in-Mechanical-Engineering/mixed_precision_for_JAX}.
 \end{abstract}
-
-\begin{document}
-\sloppy
-\maketitle
 
@@
 To extend JAX's capabilities, several libraries have been developed, including Equinox, Flax, and Optax~\citep{kidger2021equinox,flax2020github,deepmind2020jax}. 
-Despite these advancements, one critical area where JAX falls short is in supporting mixed-precision training, a method essential for efficiently handling large models~\cite{mixed_precision_paper}. 
+Despite these advancements, one critical area where JAX falls short is in supporting mixed-precision training, a method essential for efficiently handling large models~\citep{mixed_precision_paper}. 
 Although JMP~\citep{jmp} was previously available for this purpose, it is no longer maintained and cannot process arbitrary JAX PyTrees commonly used in modern neural-network toolboxes.
@@
 Mixed-precision training plays a pivotal role in optimizing the performance of large model training by executing both forward and backward passes in half-precision floating-point format (\qty{16}{\bit}).
-This technique significantly reduces memory requirements for intermediate activations and speeds up training by minimizing memory transfer overheads while utilizing specialized half-precision tensor cores (when supported by hardware), all without compromising accuracy~\cite{mixed_precision_paper}.
+This technique significantly reduces memory requirements for intermediate activations and speeds up training by minimizing memory transfer overheads while utilizing specialized half-precision tensor cores (when supported by hardware), all without compromising accuracy~\citep{mixed_precision_paper}.
 
@@
 It further introduces gradient transformations and optimizer wrappers that enable seamless conversion from a full-precision pipeline to mixed precision with minimal modifications.
-The source code, documentation, and usage examples for \mpx{} are accessible at \url{https://github.com/AlexGraefe/mixed_precision_for_JAX}.
+The source code, documentation, and usage examples for \mpx{} are accessible at \url{github.com/Data-Science-in-Mechanical-Engineering/mixed_precision_for_JAX}.
 
@@
 The \mpx{} library offers essential transformations for mixed precision training while preserving JAX's low-level flexibility. 
-Similar to JMP~\cite{jmp}, implementationts design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice.
+Similar to JMP~\citep{jmp}, implementationts design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice.
 Consequently, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision.
 
-Specifically, \mpx{} first defines functions for casting PyTrees (Section\ref{sec:castingpytrees}), then utilizes these functions to cast both general functions (Section\ref{sec:castingfunctions}) and gradient calculations (Section\ref{sec:castinggradients}) into mixed precision.
+Specifically, \mpx{} first defines functions for casting PyTrees (Section~\ref{sec:castingpytrees}), then utilizes these functions to cast both general functions (Section~\ref{sec:castingfunctions}) and gradient calculations (Section~\ref{sec:castinggradients}) into mixed precision.
 The gradient calculations include dynamic loss scaling (Section~\ref{sec:scaling}) to stabilize mixed precision training.
@@
 
-Since \mpx{} produces gradients as PyTrees, they are directly compatible with Optax optimizers~\cite{deepmind2020jax}.
+Since \mpx{} produces gradients as PyTrees, they are directly compatible with Optax optimizers~\citep{deepmind2020jax}.
 However, as previously discussed, it may be necessary to skip optimizer updates if the gradients become infinite due to loss scaling. 
@@
 
-\begin{example}[h]
-    \centering
-    \begin{minipage}{0.95\linewidth}
-    \begin{mdframed}[
-    backgroundcolor=white!10!black,
-    leftmargin=0.5cm,hidealllines=true,%
-  innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
-]
-        % \fontsize{8pt}{8pt}\selectfont
-            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+\begin{example}[t]
+\begin{pbox}[]{}
+\begin{minted}[linenos, breaklines, mathescape, fontsize=\footnotesize, xleftmargin=2em]{python}
 class MultiHeadAttentionBlock(eqx.Module):
@@
         attention_scores = q @ k.T / jnp.sqrt(q.shape[-1])
-        <@\textcolor{white}{attention\_scores = mpx.force\_full\_precision(}@>
-            <@\textcolor{white}{jax.nn.softmax, attention\_scores.dtype)(attention\_scores, axis=-1)}@>
+        attention_scores = mpx.force_full_precision(
+            jax.nn.softmax, attention_scores.dtype)(attention_scores, axis=-1)
         return attention_scores @ v
@@
     def __call__(self, inputs):
-        <@\textcolor{white}{inputs\_after\_layernorm = jax.vmap(mpx.force\_full\_precision(}@>
-            <@\textcolor{white}{self.layer\_norm, inputs.dtype))(inputs)}@>
+        inputs_after_layernorm = jax.vmap(mpx.force_full_precision(
+            self.layer_norm, inputs.dtype))(inputs)
         qs = jax.vmap(self.dense_qs)(inputs_after_layernorm)
@@
         return outputs
-        \end{lstlisting}
-        \end{mdframed}
-    \end{minipage}
+\end{minted}
+\end{pbox}
 \caption{Implementation of multi-headed self-attention via MPX.
-\capt{For the largest part, the model can be implemented in common toolboxes (here, Equinox~\cite{kidger2021equinox}).
+\capt{For the largest part, the model can be implemented in common toolboxes (here, Equinox~\citep{kidger2021equinox}).
 Only critical function that should be implemented in full precision, like softmax or layernorm (due to mean and std calculations), have to be transformed via \mpx{}.}}
@@
         \centering
-        \begin{mdframed}[
-            backgroundcolor=white!10!black,
-            leftmargin=0.5cm,hidealllines=true,%
-            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
-        ]
-            % \fontsize{8pt}{8pt}\selectfont
-            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+        \begin{pbox}[]{}
+            \begin{minted}[linenos, breaklines, mathescape, fontsize=\footnotesize, xleftmargin=2em]{python}
 grads = eqx.filter_grad(loss)(model, batch)
@@
 model = eqx.apply_updates(model, updates)
-            \end{lstlisting}
-        \end{mdframed}
-        \vspace{-0.5cm}
+        \end{minted}
+    \end{pbox}
         \caption{Full Precision via Equinox}
-        \vspace{0.5cm}
     \end{subfigure}\hspace{0.1cm}
@@
         \centering
-        \begin{mdframed}[
-            backgroundcolor=white!10!black,
-            leftmargin=0.5cm,hidealllines=true,%
-            innerleftmargin=0.2cm,innerrightmargin=0.2cm,innertopmargin=0cm,innerbottommargin=-0.10cm
-        ]
-            % \fontsize{8pt}{8pt}\selectfont
-            \begin{lstlisting}[, escapeinside={<@}{@>},language=Python, basicstyle=\ttfamily\linespread{1.1}\footnotesize\color{white!70!black}, frame=l, label={lst:model}, keywordstyle=\color{blue!30!white}, commentstyle=\color{gray}, stringstyle=\color{orange!80!black}, numbers=left, backgroundcolor=\color{white!10!black}, numbers=left, framesep=0.8cm, framerule=0pt,  xleftmargin=0.7cm,]
+        \begin{pbox}[]{}
+            \begin{minted}[linenos, breaklines, mathescape, fontsize=\footnotesize, xleftmargin=2em]{python}
 loss_scaling, grads_finite, grads = mpx.filter_grad(loss, loss_scaling)(
@@
     model, optimizer, optimizer_state, grads,grads_finite)
-            \end{lstlisting}
-        \end{mdframed}
-        \vspace{-0.5cm}
-        \caption{mixed precision via MPX}
+            \end{minted}
+            \end{pbox}
+            \caption{Mixed Precision via \mpx{}}
     \end{subfigure}
@@
 
-We train vision transformers (ViT)~\cite{dosovitskiy2020image,steiner2021train}.
+We train vision transformers (ViT)~\citep{dosovitskiy2020image,steiner2021train}.
 On the desktop PC, the transformer features a size of 256 with residual blocks containing one hidden layer of 800 neurons.
@@
 For training speed assessment, both settings are turned off.
+Due to technical reasons, we only measure VRAM consumption on the desktop PC as an accurate VRAM measurement was not possible on the high-performance cluster.
 
@@
     \centering
-    \begin{subfigure}[t]{0.49\linewidth}
+    \begin{subfigure}[t]{0.9\linewidth}
     \begin{tikzpicture}
@@
             width=0.8\textwidth,
-            height=7cm,
+            height=5cm,
             xlabel={Batch Size},
@@
     \end{tikzpicture}
-    \caption{Desktop Computer (RTX4070)}
+    %\caption{Desktop Computer (RTX4070)}
     \end{subfigure}
-    \begin{subfigure}[t]{0.49\linewidth}
-    \begin{tikzpicture}
-        \begin{axis}[
-            width=0.8\textwidth,
-            height=7cm,
-            xlabel={Batch Size},
-            ylabel={VRAM (\qty{}{\giga\byte})},
-            legend pos=south east,
-            grid=major,
-            grid style={dashed,gray!30},
-            xmin=0,
-            ymin=0
-        ]
-        % First graph: Full Precision
-        \addplot[
-            color=blue,
-            mark=o,
-            thick
-        ] coordinates {
-            (128, 2.290)
-            (256, 4.724)
-            (512, 8.860)
-        };
-        \addlegendentry{Full Precision}
+    % \begin{subfigure}[t]{0.49\linewidth}
+    % \begin{tikzpicture}
+    %     \begin{axis}[
+    %         width=0.8\textwidth,
+    %         height=7cm,
+    %         xlabel={Batch Size},
+    %         ylabel={VRAM (\qty{}{\giga\byte})},
+    %         legend pos=south east,
+    %         grid=major,
+    %         grid style={dashed,gray!30},
+    %         xmin=0,
+    %         ymin=0
+    %     ]
+    %     % First graph: Full Precision
+    %     \addplot[
+    %         color=blue,
+    %         mark=o,
+    %         thick
+    %     ] coordinates {
+    %         (128, 2.290)
+    %         (256, 4.724)
+    %         (512, 8.860)
+    %     };
+    %     \addlegendentry{Full Precision}
 
-        % Second graph: Mixed Precision
-        \addplot[
-            color=red,
-            mark=square*,
-            thick,
-            dashed
-        ] coordinates {
-            (128, 1.52)
-            (256, 2.418)
-            (512, 4.900)
-            (768, 7.054)
-            (1024, 9.222)
-        };
-        \addlegendentry{Mixed Precision}
-        \end{axis}
-    \end{tikzpicture}
-    \caption{High-Performance Cluster ($4\times$ H100)}
-    \end{subfigure}
-    \caption{Comparison of GPU VRAM consumed for full precision and mixed precision as a function of the number of batches.}
+    %     % Second graph: Mixed Precision
+    %     \addplot[
+    %         color=red,
+    %         mark=square*,
+    %         thick,
+    %         dashed
+    %     ] coordinates {
+    %         (128, 1.52)
+    %         (256, 2.418)
+    %         (512, 4.900)
+    %         (768, 7.054)
+    %         (1024, 9.222)
+    %     };
+    %     \addlegendentry{Mixed Precision}
+    %     \end{axis}
+    % \end{tikzpicture}
+    % \caption{High-Performance Cluster ($4\times$ H100)}
+    % \end{subfigure}
+    \caption{Comparison of GPU VRAM consumed for full precision and mixed precision as a function of the number of batches on the desktop PC.}
     \label{fig:memory}
@@
 The speedup is caused by reduced memory loading times as the RTX4070 streaming-multiprocessors feature no computing speedup for half precision.
-The cluster features a similar behavior, where both RAM and processing times are reduced.
-RAM is reduced by factor $1.8\times$ and processing time by up to $1.57\times$.
+The cluster features a similar behavior, where processing time is reduced by up to $1.57\times$.
 
@@
 \mpx{} offers an easy-to-use solution for mixed-precision training in JAX.
-It is able to handling arbitrary PyTrees and provides a seamless integration with popular libraries like Equinox and Flax. 
-Through dynamic loss scaling and precision casting, MPX ensures efficient memory usage and accelerates training without compromising model accuracy. 
-The evaluation demonstrates significant reductions in RAM usage and processing times, particularly on hardware that supports half-precision computation. 
-Overall, MPX presents a compelling advancement in optimizing large-scale neural network training pipelines while maintaining flexibility and ease of use.
+It is able to handle arbitrary PyTrees and provides a seamless integration with popular libraries like Equinox and Flax. 
+At its core, \mpx{} allows casting functions to mixed precision and provides wrappers that allow to calculate gradients using mixed precision with dynamic loss scaling.
+The evaluation demonstrates significant reductions in RAM usage and processing times. 
+Overall, MPX presents a compelling toolbox for optimizing large-scale neural network training pipelines while maintaining flexibility and ease of use.
 
 \section{Acknowledgements}
-We express our gratitude to Patrick Kidger for Equinox and Google DeepMind for JMP, which served as the foundation for this implementation.
+We thank Patrick Kidger for providing Equinox and Google DeepMind for JMP, which served as the foundation for this implementation.
 
@@
 
-% \begin{thebibliography}{9}
-% \bibitem{jmp} 
-% JMP: JAX mixed precision
-% \newblock \url{https://github.com/google-deepmind/jmp}
-
-% \bibitem{equinox}
-% Equinox: Neural Networks in JAX
-% \newblock \url{https://docs.kidger.site/equinox/}
-
-% \bibitem{nvidia_amp}
-% NVIDIA Automatic mixed precision
-% \newblock \url{https://developer.nvidia.com/automatic-mixed precision}
-
-% \bibitem{mixed_precision_paper}
-% P. Micikevicius et al.
-% \newblock ``mixed precision Training''
-% \newblock arXiv:1710.03740, 2017
-% \end{thebibliography}
-
 \end{document}

### 2025-07-03 docs/paper/main.tex
major refactoring of docs and namespaces

--- 2025-06-10 doc/paper/main.tex
@@
 % \date{}
+
+% # build: python3 -m build --sdist
+% # upload: twine upload --repository pypi dist/*
 
@@
 The \mpx{} library offers essential transformations for mixed precision training while preserving JAX's low-level flexibility. 
-Similar to JMP~\citep{jmp}, implementationts design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice.
+Similar to JMP~\citep{jmp}, our implementation's design leverages JAX's type-promotion behavior\footnote{\url{https://docs.jax.dev/en/latest/jep/9407-type-promotion.html}}, ensuring that operations are executed in the precision to which inputs and outputs have been cast—provided all constants in the function are positioned on the left side of the type-promotion lattice.
 Consequently, \mpx{} primarily focuses on casting function inputs and outputs to the desired precision.
@@
 The desktop PC trains the ViT on CIFAR100 while the high-performance cluster handles ImageNet1k, dividing each batch equally across GPUs using a data-parallel approach.
-We train without gradient accumulation, which could further optimize runtime and memory consumption
+We train without gradient accumulation, which could further optimize runtime and memory consumption.
 
@@
 
+We thank Lukas Wildberger for helpful discussions and suggestions.
+
 The authors gratefully acknowledge the computing time provided to them at the NHR Center NHR4CES at RWTH Aachen University (project number p0021919).

### 2025-07-03 docs/conf.py

@@ -1,0 +1,35 @@
+# Configuration file for the Sphinx documentation builder.
+#
+# For the full list of built-in configuration values, see the documentation:
+# https://www.sphinx-doc.org/en/master/usage/configuration.html
+
+# -- Project information -----------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#project-information
+
+import sys
+from pathlib import Path
+
+sys.path.insert(0, str(Path('..').resolve()))
+
+project = 'MPX'
+copyright = '2025, Alexander Gräfe'
+author = 'Alexander Gräfe'
+
+# -- General configuration ---------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#general-configuration
+
+extensions = ['sphinx.ext.autodoc']
+
+templates_path = ['_templates']
+exclude_patterns = ['_build', 'Thumbs.db', '.DS_Store']
+
+
+
+# -- Options for HTML output -------------------------------------------------
+# https://www.sphinx-doc.org/en/master/usage/configuration.html#options-for-html-output
+
+html_theme = 'pydata_sphinx_theme'
+html_theme_options = {
+  "show_toc_level": 2
+}
+html_static_path = ['_static']

### 2025-07-03 mpx/__init__.py

@@ -1,13 +1,10 @@
 """
 Mixed Precision for JAX - A library for mixed precision training in JAX
-
-# build: python3 -m build --sdist
-# upload: twine upload --repository pypi dist/* 
 """
 
 __version__ = "0.1.5"
 
-from .cast import (
+from ._cast import (
     cast_tree,
     cast_to_float32,
     cast_to_float16,
@@ -15,10 +12,11 @@ from .cast import (
     cast_to_full_precision,
     cast_to_half_precision,
     force_full_precision,
+    cast_function,
 )
-from .dtypes import half_precision_datatype, set_half_precision_datatype, FLOAT16_MAX, BFLOAT16_MAX
-from .loss_scaling import DynamicLossScaling, all_finite, scaled
-from .grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update
+from ._dtypes import half_precision_datatype, set_half_precision_datatype, FLOAT16_MAX, BFLOAT16_MAX, HALF_PRECISION_DATATYPE
+from ._loss_scaling import DynamicLossScaling, all_finite, scaled
+from ._grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update, calculate_scaled_grad
 
 __all__ = [
     # Cast functions
@@ -29,6 +27,7 @@ __all__ = [
     'cast_to_full_precision',
     'cast_to_half_precision',
     'force_full_precision',
+    'cast_function',
     
     # Dtype functions
     'half_precision_datatype',
@@ -44,4 +43,5 @@ __all__ = [
     'filter_grad',
     'filter_value_and_grad',
     'optimizer_update',
+    'calculate_scaled_grad',
 ]

### 2025-07-03 mpx/_cast.py

@@ -1,30 +1,4 @@
 """
-MIT License
-
-Copyright (c) 2025 Alexander Gräfe
-
-Permission is hereby granted, free of charge, to any person obtaining a copy
-of this software and associated documentation files (the "Software"), to deal
-in the Software without restriction, including without limitation the rights
-to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-copies of the Software, and to permit persons to whom the Software is
-furnished to do so, subject to the following conditions:
-
-The above copyright notice and this permission notice shall be included in all
-copies or substantial portions of the Software.
-
-THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
-SOFTWARE.
-
-Tools for mixer precision training. Methods and general code architecture are from jmp https://github.com/google-deepmind/jmp. This can be seen as a port and extension of JMP tot equinox.
-"""
-
-"""
 Functions for casting of Pytrees.
 """
 
@@ -34,7 +8,7 @@ import equinox as eqx
 
 from jaxtyping import Array, Float, Int, PyTree, PRNGKeyArray 
 
-from .dtypes import half_precision_datatype
+from ._dtypes import half_precision_datatype
 
 def cast_tree(tree: PyTree, dtype):
     """

### 2025-07-03 mpx/_grad_tools.py

@@ -1,29 +1,3 @@
-"""
-MIT License
-
-Copyright (c) 2025 Alexander Gräfe
-
-Permission is hereby granted, free of charge, to any person obtaining a copy
-of this software and associated documentation files (the "Software"), to deal
-in the Software without restriction, including without limitation the rights
-to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-copies of the Software, and to permit persons to whom the Software is
-furnished to do so, subject to the following conditions:
-
-The above copyright notice and this permission notice shall be included in all
-copies or substantial portions of the Software.
-
-THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
-SOFTWARE.
-
-Tools for mixer precision training. Methods and general code architecture are from jmp https://github.com/google-deepmind/jmp. This can be seen as a port and extension of JMP tot equinox.
-"""
-
 """Filtering tools for mixer precision training."""
 
 
@@ -33,8 +7,8 @@ import equinox as eqx
 
 import optax
 
-import mpx.cast as cast
-import mpx.loss_scaling as loss_scaling
+import mpx._cast as cast
+import mpx._loss_scaling as loss_scaling
 
 from jaxtyping import PyTree, Bool
 

### 2025-07-03 mpx/_loss_scaling.py

@@ -1,29 +1,3 @@
-"""
-MIT License
-
-Copyright (c) 2025 Alexander Gräfe
-
-Permission is hereby granted, free of charge, to any person obtaining a copy
-of this software and associated documentation files (the "Software"), to deal
-in the Software without restriction, including without limitation the rights
-to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-copies of the Software, and to permit persons to whom the Software is
-furnished to do so, subject to the following conditions:
-
-The above copyright notice and this permission notice shall be included in all
-copies or substantial portions of the Software.
-
-THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
-SOFTWARE.
-
-Tools for mixer precision training. Methods and general code architecture are from jmp https://github.com/google-deepmind/jmp. This can be seen as a port and extension of JMP tot equinox.
-"""
-
 """Tools for automatic loss scaling (https://docs.nvidia.com/deeplearning/performance/mixed-precision-training/index.html)."""
 
 
@@ -60,6 +34,23 @@ def all_finite(tree: PyTree) -> Array:
 
 
 def scaled(func: callable, scaling: 'DynamicLossScaling', has_aux: bool = False) -> callable:
+    """
+    Scales the output of a function using dynamic loss scaling.
+    This decorator wraps a given function such that its output is scaled using the
+    provided dynamic loss scaling object. If the wrapped function returns auxiliary
+    data (indicated by has_aux=True), only the primary value is scaled; otherwise, the
+    sole returned value is scaled.
+    Parameters:
+        func (callable): The original function whose output is to be scaled.
+        scaling (DynamicLossScaling): An object providing a `scale` method for scaling
+            the function's output.
[…]

### 2025-07-03 tests/test_cast.py

@@ -5,7 +5,7 @@ import equinox as eqx
 from jaxtyping import Array, Float, Int, PyTree
 import numpy as np
 
-from mpx.cast import (
+from mpx import (
     cast_tree,
     cast_to_float32,
     cast_to_float16,
@@ -15,7 +15,7 @@ from mpx.cast import (
     force_full_precision,
     cast_function,
 )
-from mpx.dtypes import HALF_PRECISION_DATATYPE
+from mpx import HALF_PRECISION_DATATYPE
 
 
 class EQXModuleBase(eqx.Module):

### 2025-07-03 tests/test_dtypes.py

@@ -1,6 +1,6 @@
 import unittest
 import jax.numpy as jnp
-from mpx.dtypes import half_precision_datatype, set_half_precision_datatype
+from mpx import half_precision_datatype, set_half_precision_datatype
 
 class TestDtypes(unittest.TestCase):
     def test_default_half_precision(self):

### 2025-07-03 tests/test_grad_tools.py

@@ -3,8 +3,8 @@ import jax
 import jax.numpy as jnp
 import equinox as eqx
 import optax
-from mpx.grad_tools import select_tree, filter_grad, filter_value_and_grad, optimizer_update, calculate_scaled_grad
-from mpx.loss_scaling import DynamicLossScaling
+from mpx import select_tree, filter_grad, filter_value_and_grad, optimizer_update, calculate_scaled_grad
+from mpx import DynamicLossScaling
 
 
 # Create a simple model for testing

### 2025-07-03 tests/test_loss_scaling.py

@@ -2,7 +2,7 @@ import unittest
 import jax
 import jax.numpy as jnp
 import equinox as eqx
-from mpx.loss_scaling import DynamicLossScaling, all_finite, scaled
+from mpx import DynamicLossScaling, all_finite, scaled
 
 class TestLossScaling(unittest.TestCase):
     def setUp(self):

### 2025-07-03 docs/conf.py
cleanup

@@ -9,7 +9,7 @@
 import sys
 from pathlib import Path
 
-sys.path.insert(0, str(Path('..').resolve()))
+# sys.path.insert(0, str(Path('..').resolve()))
 
 project = 'MPX'
 copyright = '2025, Alexander Gräfe'

### 2025-07-03 docs/conf.py
changed workflow again

@@ -9,7 +9,7 @@
 import sys
 from pathlib import Path
 
-# sys.path.insert(0, str(Path('..').resolve()))
+sys.path.insert(0, str(Path('..').resolve()))
 
 project = 'MPX'
 copyright = '2025, Alexander Gräfe'

### 2025-07-03 mpx/__init__.py
updated readme

@@ -2,7 +2,7 @@
 Mixed Precision for JAX - A library for mixed precision training in JAX
 """
 
-__version__ = "0.1.5"
+__version__ = "0.1.6"
 
 from ._cast import (
     cast_tree,

### 2025-10-25 examples/train_vit.py
updated readme and examples to reflect that many operations do not need to be forced to full precision as they already do this internally

@@ -179,7 +179,6 @@ def train_epoch(model: eqx.Module,
         if counter > 0:
             training_times.append(time.time() - start_time)
         counter += 1
-        print(time.time() - start_time)
         loss_value += loss_batch.astype(jnp.float32) * len(batch["input"])
         num_datapoints += len(batch["input"])
     

### 2025-10-25 examples/transformer.py

@@ -41,7 +41,7 @@ class ResidualBlock(eqx.Module):
     i.e. y = LayerNorm(MLP(x) + Residual(x)), where Resiudual(x) = x if transform_residual is False and Residual(x) = DenseLayer(x) if transform_residual is True
     """
 
-    layers: list[DenseLayer]
+    layers: list[eqx.nn.Linear]
     residual_layer: DenseLayer | None
     dropout: eqx.nn.Dropout
 
@@ -65,23 +65,23 @@ class ResidualBlock(eqx.Module):
         layers = []
         pruning_masks = []
         key, subkey = jax.random.split(key)
-        layers.append(DenseLayer(input_dim, feature_dim, subkey))
+        layers.append(eqx.nn.Linear(input_dim, feature_dim, key=subkey))
 
         if num_layers >= 2:
             for _ in range(num_layers - 2):
                 key, subkey = jax.random.split(key)
-                layers.append(DenseLayer(feature_dim, 
+                layers.append(eqx.nn.Linear(feature_dim, 
                                     feature_dim,
-                                    subkey))
+                                    key=subkey))
             
             key, subkey = jax.random.split(key)
-            layers.append(DenseLayer(feature_dim, output_dim, subkey))
+            layers.append(eqx.nn.Linear(feature_dim, output_dim, key=subkey))
         
         self.layers = layers
 
         assert transform_residual or input_dim == output_dim, "input_dim must be equal to output_dim if transform_residual is False"
         if transform_residual:
-            self.residual_layer = DenseLayer(input_dim, output_dim, key)
+            self.residual_layer = eqx.nn.Linear(input_dim, output_dim, key=key)
         else:
             self.residual_layer = None
 
@@ -110,7 +110,6 @@ class ResidualBlock(eqx.Module):
             outputs = self.activation_fn(outputs)
 
             for i, layer in enumerate(self.layers[1:-1]):
-                key, subkey = jax.random.split(key)
                 outputs = self.activation_fn(layer(outputs))
 
             outputs = self.layers[-1](outputs)
@@ -128,11 +127,7 @@ class ResidualBlock(eqx.Module):
     
 
 class MultiHeadAttentionBlock(eqx.Module):
-    dense_qs: DenseLayer 
-    dense_ks: DenseLayer
-    dense_vs: DenseLayer
-
-    dense_o: DenseLayer
+    attention: eqx.nn.MultiheadAttention
 
[…]

### 2025-10-26 docs/paper/main.tex
updated readme and examples to reflect that many operations do not need to be forced to full precision as they already do this internally
added a small playground script for fp8 computations

@@
-\documentclass[copyright, logo]{dsme}
+\documentclass[logo]{dsme}
+% \documentclass[twocolumn]{article}
 
@@
 Furthermore, \mpx{} includes \texttt{force\_full\_precision(func, return\_dtype)}, which ensures computations are carried out in full precision. 
-This feature is particularly important for operations that are prone to overflow in \texttt{float16}, such as \texttt{sum}, \texttt{mean}, or \texttt{softmax}.
+This feature is particularly important for operations that are prone to overflow in \texttt{float16}.
 
@@
 
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}[t]{0.99\textwidth}
-        \centering
-        \includegraphics[width=0.99\textwidth]{assets/EQXOverview.pdf}
-        \caption{Full Precision Training via Equinox}
-        \vspace{0.2cm}
-    \end{subfigure}\hspace{0.1cm}
-    \begin{subfigure}[t]{0.99\textwidth}
-        \centering
-        \includegraphics[width=0.99\textwidth]{assets/MPXOverview.pdf}
-        \caption{Half Precision Training via Equinox and \mpx{}}
-    \end{subfigure}\hspace{0.1cm}
-    \caption{Transformation of training pipelines via MPX.
+% \begin{figure}[t]
+%     \centering
+%     \begin{subfigure}[t]{0.99\textwidth}
+%         \centering
+%         \includegraphics[width=0.99\textwidth]{assets/EQXOverview.pdf}
+%         \caption{Full Precision Training via Equinox}
+%         \vspace{0.2cm}
+%     \end{subfigure}\hspace{0.1cm}
+%     \begin{subfigure}[t]{0.99\textwidth}
+%         \centering
+%         \includegraphics[width=0.99\textwidth]{assets/MPXOverview.pdf}
+%         \caption{Half Precision Training via Equinox and \mpx{}}
+%     \end{subfigure}\hspace{0.1cm}
+%     \caption{Transformation of training pipelines via MPX.
 \capt{An existing training pipeline (Top figure) can be transformed, by applying \texttt{@mpx.filter\_grad} to the forward pass.
 It automatically casts all inputs to half precision and calculates the gradient via dynamic loss scaling and \texttt{eqx.filter\_grad}.}}
-    \label{fig:overview}
-\end{figure}
+%     \label{fig:overview}
+% \end{figure}
 
@@
 \subsection{Model Implementation}
+\label{sec:eqxvsmpxmodelupdate}
 
-\begin{example}[t]
-\begin{pbox}[]{}
-\begin{minted}[linenos, breaklines, mathescape, fontsize=\footnotesize, xleftmargin=2em]{python}
-class MultiHeadAttentionBlock(eqx.Module):
-    dense_qs: eqx.nn.Linear
-    dense_ks: eqx.nn.Linear
-    dense_vs: eqx.nn.Linear
-    dense_o: eqx.nn.Linear
-    num_heads: int
-    layer_norm: eqx.nn.LayerNorm
+% \begin{example}[t]
+% \begin{pbox}[]{}
+% \begin{minted}[linenos, breaklines, mathescape, fontsize=\footnotesize, xleftmargin=2em]{python}
+% class MultiHeadAttentionBlock(eqx.Module):
+%     dense_qs: eqx.nn.Linear
+%     dense_ks: eqx.nn.Linear
+%     dense_vs: eqx.nn.Linear
+%     dense_o: eqx.nn.Linear
+%     num_heads: int
+%     layer_norm: eqx.nn.LayerNorm
 
-    def __init__(self, feature_dim, num_heads, key):
-        self.num_heads = num_heads
-        key, subkey = jax.random.split(key)
-        self.dense_qs = eqx.nn.Linear(
-            feature_dim, feature_dim, key=subkey)
-        # same for dense_ks, dense_vs, dense_o
+%     def __init__(self, feature_dim, num_heads, key):
+%         self.num_heads = num_heads
+%         key, subkey = jax.random.split(key)
+%         self.dense_qs = eqx.nn.Linear(
+%             feature_dim, feature_dim, key=subkey)
+%         # same for dense_ks, dense_vs, dense_o
 
-        self.layer_norm = eqx.nn.LayerNorm(feature_dim)
+%         self.layer_norm = eqx.nn.LayerNorm(feature_dim)
 
-    def attention(q, k, v):
-        attention_scores = q @ k.T / jnp.sqrt(q.shape[-1])
-        attention_scores = mpx.force_full_precision(
-            jax.nn.softmax, attention_scores.dtype)(attention_scores, axis=-1)
-        return attention_scores @ v
+%     def attention(q, k, v):
+%         attention_scores = q @ k.T / jnp.sqrt(q.shape[-1])
+%         attention_scores = mpx.force_full_precision(
+%             jax.nn.softmax, attention_scores.dtype)(attention_scores, axis=-1)
+%         return attention_scores @ v
 
-    def __call__(self, inputs):
-        inputs_after_layernorm = jax.vmap(mpx.force_full_precision(
-            self.layer_norm, inputs.dtype))(inputs)
-        qs = jax.vmap(self.dense_qs)(inputs_after_layernorm)
-        qs = es.jax_einshape("n(hf)->hnf", qs, h=self.num_heads)
-        # same for ks and vs...
+%     def __call__(self, inputs):
+%         inputs_after_layernorm = jax.vmap(mpx.force_full_precision(
+%             self.layer_norm, inputs.dtype))(inputs)
+%         qs = jax.vmap(self.dense_qs)(inputs_after_layernorm)
+%         qs = es.jax_einshape("n(hf)->hnf", qs, h=self.num_heads)
+%         # same for ks and vs...
 
-        outputs = jax.vmap(self.attention, in_axes=(0, 0, 0))(qs, ks, vs)
-        outputs = es.jax_einshape("hnf->n(hf)", outputs)
-        outputs = jax.vmap(self.dense_o)(outputs)
-        outputs += inputs
+%         outputs = jax.vmap(self.attention, in_axes=(0, 0, 0))(qs, ks, vs)
+%         outputs = es.jax_einshape("hnf->n(hf)", outputs)
+%         outputs = jax.vmap(self.dense_o)(outputs)
+%         outputs += inputs
 
-        return outputs
-\end{minted}
-\end{pbox}
-\caption{Implementation of multi-headed self-attention via MPX.
+%         return outputs
+% \end{minted}
+% \end{pbox}
+% \caption{Implementation of multi-headed self-attention via MPX.
 \capt{For the largest part, the model can be implemented in common toolboxes (here, Equinox~\citep{kidger2021equinox}).
 Only critical function that should be implemented in full precision, like softmax or layernorm (due to mean and std calculations), have to be transformed via \mpx{}.}}
-\label{ex:model}
-\end{example}
+% \label{ex:model}
+% \end{example}
 
@@
 As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\citep{flax2020github} or Equinox~\citep{kidger2021equinox} can be used in conjunction with \mpx{} (Example~\ref{ex:model}).
-Only parts of the model that contain critical operations with a high risk of overflows, like sum or mean, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.
+Only parts of the model that contain critical operations with a high risk of overflows, must be transformed to full precision via \texttt{mpx.force\_full\_precision}.
+However, many jax and also equinox already implement many critical functions in a numerically stable way by casting to full precision if necessary, so that only few changes are necessary.
 
-\subsection{Model Updates}
-\label{sec:eqxvsmpxmodelupdate}
+% \subsection{Model Updates}
+% \label{sec:eqxvsmpxmodelupdate}
 Model updates via \mpx{} closely resemble those performed with Equinox (Example~\ref{ex:update}). 

### 2025-10-26 mpx/experimental/fp8.py

@@ -1,0 +1,111 @@
+from functools import partial
+
+import jax
+import jax.numpy as jnp
+import equinox as eqx
+
+from collections.abc import Sequence
+from jaxtyping import ArrayLike  
+
+import quax
+
+def max_val(dtype):
+    return (jnp.finfo(dtype).max).astype(jnp.float32)
+
+@partial(jax.custom_vjp, nondiff_argnames=("dtype8", 'dimension_numbers', 'precision', 'preferred_element_type', 'out_sharding'))
+def quantized_multiplication(a: ArrayLike, b: ArrayLike, dtype8, dimension_numbers, precision, preferred_element_type, out_sharding):
+    a_max = jnp.max(jnp.abs(a))
+    b_max = jnp.max(jnp.abs(b))
+    max_dtype = max_val(dtype8)
+    scaling_a = max_dtype / (a_max + 1e-8)
+    scaling_b = max_dtype / (b_max + 1e-8)
+
+    a_q = (a * scaling_a).astype(dtype8)
+    b_q = (b * scaling_b).astype(dtype8)
+
+    result_q = jax.lax.dot_general_p.bind(a_q, b_q, dimension_numbers=dimension_numbers, precision=precision, preferred_element_type=preferred_element_type, out_sharding=out_sharding)
+
+    result = (result_q.astype(jnp.float32)) / (scaling_a * scaling_b)
+    return result
+
+
+def quantized_multiplication_fwd(a: ArrayLike, b: ArrayLike, dtype8, dimension_numbers, precision, preferred_element_type, out_sharding):
+    a_max = jnp.max(jnp.abs(a))
+    b_max = jnp.max(jnp.abs(b))
+    max_dtype = max_val(dtype8)
+    scaling_a = max_dtype / (a_max + 1e-8)
+    scaling_b = max_dtype / (b_max + 1e-8)
+
+    a_q = (a * scaling_a).astype(dtype8)
+    b_q = (b * scaling_b).astype(dtype8)
+    # we want to save the quantized versions for the backward pass to save memory
+    return quantized_multiplication(a, b, dtype8, dimension_numbers, precision, preferred_element_type, out_sharding), (a_q, b_q, scaling_a, scaling_b)
+
+# f_bwd :: (c, CT b) -> CT a
[…]

### 2025-10-26 mpx/experimental/fp8.py
Removed debug prints

@@ -43,15 +43,10 @@ def quantized_multiplication_fwd(a: ArrayLike, b: ArrayLike, dtype8, dimension_n
 
 # f_bwd :: (c, CT b) -> CT a
 def quantized_multiplication_bwd(dtype8, dimension_numbers, precision, preferred_element_type, out_sharding, c, dy_dc):
-  print("--------")
-  print(c)
-  print(dy_dc)
   a_q, b_q, scaling_a, scaling_b = c
   # backward is performed in fp32 TODO allow to change it.
   a = a_q.astype(jnp.float32) / scaling_a
   b = b_q.astype(jnp.float32) / scaling_b
-  print(dy_dc.shape)
-  print(b.shape)
   dy_da = jax.lax.dot_general_p.bind(dy_dc, b.T, dimension_numbers=dimension_numbers, precision=precision, preferred_element_type=preferred_element_type, out_sharding=out_sharding)
   dy_db = jax.lax.dot_general_p.bind(a.T, dy_dc, dimension_numbers=dimension_numbers, precision=precision, preferred_element_type=preferred_element_type, out_sharding=out_sharding)
 
@@ -61,8 +56,6 @@ quantized_multiplication.defvjp(quantized_multiplication_fwd, quantized_multipli
 
 @quax.register(jax.lax.dot_general_p)
 def _(lhs: ArrayLike, rhs: ArrayLike, **params):
-    print("Performing a matmul!")
-    print(params)
     return quantized_multiplication(lhs, rhs, jnp.float8_e4m3, **params)
 
 if __name__ == "__main__":
